---
title: "Wood-Poppy Integral Projection Model"
author: "Amy E. Wiedenfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

Integral Projection Model for the wood-poppy

Code from:
https://cmerow.github.io/RDataScience/21_Intro_IPMs.html
https://github.com/jloesberg/Primula/blob/main/scripts/primula_IPM.rmd

-along with other papers, mentioned below

# Load Data and Packages
```{r message=FALSE, warning=FALSE}
#libraries
library(tidyverse)
library(lubridate)
library(stringr)
library(nlme)
library(popbio)
library(cowplot)
library(gridExtra)
library(popdemo)
library(ggpubr)
library(arm)
library(car)
library(fields)

#Data

# Formatted for the IPM
SurvData <- read.csv("~/Research/Data/Github Code/Stylophorum-IPMs/SurvData_IProjM.csv")
```


Add data on average number of seeds by population
```{r}
#Average seed count by population (from Emma Neigel's 2022 data)
Bayfield_seeds <- 38.6
Fanshawe_seeds <- 87.5
Ilderton_seeds <- 37.0
Meadowlily_seeds <- 21.3

#no data on RBG, and no seeds collected at Denfield
#For these two locations, I will average the four sets
# of seed averages that I have
Denfield_seeds <- mean(Bayfield_seeds, Fanshawe_seeds,
                       Ilderton_seeds, Meadowlily_seeds)
RBG_seeds <- mean(Bayfield_seeds, Fanshawe_seeds,
                  Ilderton_seeds, Meadowlily_seeds)
```

# Model Form

The first step in constructing the IPM is to construct a life-cycle
 diagram for the species, and then, based on that, construct
 the model equation. The model equation (or equations) determine
 what vital rate regressions and probabilities to calculate, and
 will later determine the form of the model.
 
 These are the equations (not written clearly--need special
  characters to do this right) for the wood-poppy.

**Number of 2-year old seeds**

S2(t+1) = (1 - probability of germination[S1] - probability of 
 germination[s2]) * S1
 
$S_{2} = (1 - g_{0})S_{1}(t)$

**Number of 1-year old seeds**

S1(t+1) = (1 - probability of germination) * integral(probability
 of flowering * number of seeds per plant)
 
$S_{1}(t + 1) = (1 - g_{0})\int\limits_L^U p_{f}(x)f_{s}(x)n(x, t)dx$

**Number of seedlings**

C(t+1) = (seed germination) * integral(probability of flowering *
 number of seeds per plant) + probability of germination * S1 +
 probability of germination * S2
 
$C(t+1) = p_{est}[g_{0}]\int\limits_L^U p_{f}(x)f_{s}(x)n(x, t)dx + g_{1}S_{1}(t) + g_{1}S_{2}(t)$

 
**Number of Established Plants (not seedlings)**
 
 n(y, t+1) =  C(t) * probability of establishment * distribution
  of recruit sizes + integral ( probability of survival * 
  probability of growth * number of individuals at time t)
  
$n(y, t+1) = C(t)p_{sc}f{sd}(y) + \int\limits_L^U p_{s}(x)g(y, x)n(x, t) dx$ 
  
## Vital Rate Regressions

The next step in the model is to determine which vital rate regressions
 and probabilities will need to be calculated, based on the model
 equation.

Probabilities
Probability of germination

Regressions
Probability of flowering
Probability of survival
probability of establishment
distribution of recruit sizes
probability of survival
probability of growth


Code info from: Merow et al. 2014, Rees et al. 2014


Models are confirmed as meeting assumptions in the IProj_M_regressions
 R script. Each regression that is selected via an anova is from the
 IProj_M_regressions R script. For models where multiple slopes and
 intercepts are used, these are also extracted just below the model.


```{r Format and Organize Data}
######### Modify Dataframes ################################

# Remove Ilderton 334 - these data are wrong
SurvData <- SurvData %>%
  filter(Individual != "334")

#Add columns defining established plants (not seedlings), seedlings,
# and last year's seedlings
Estab_Plants <- SurvData %>%
  #seedling to adult is discrete transition
  filter(Recruit_t == 0) %>%
  #I don't want seedlings in this dataset
  filter(Recruit_t1 == 0)

#For this, I only want seedlings that germinate in t+1
Recruits <- SurvData %>%
  filter(Recruit_t == 0) %>%
  filter(Recruit_t1 == 1)

# I want Year to be numeric, but just as 1,2, etc.
# Year "1" is 2022
Estab_Plants$Year_t <- as.factor(Estab_Plants$Year_t)
Estab_Plants$Year_t <- as.numeric(Estab_Plants$Year_t)
# Make sure that Population is also a factor
Estab_Plants$Population <- as.factor(Estab_Plants$Population)

#For the seeds per plant
#This makes it so that it's only plants that have at least one seed
Estab_Plants2 <- Estab_Plants %>%
  filter(!is.na(Height_cm_t)) %>%
  filter(!is.na(Seed_Number_t1)) %>%
  filter(Seed_Number_t1 != 0)

# For the GLM, I need the log of seed numbers for the model to work
# well. I'll convert it back in the functions later
Estab_Plants2$Seed_Number_t_x <- log(Estab_Plants2$Seed_Number_t)
Estab_Plants2$Seed_Number_t1_x <- log(Estab_Plants2$Seed_Number_t1)

#For Flowering probability
#make flowering binary
Estab_Plants <- Estab_Plants %>%
  mutate(Flowering = case_when(Flower_Number_t1 > 0 ~ 1,
                               Flower_Number_t1 == 0 ~ 0,
                               TRUE ~ 0))
 
# Split by population
# Because I can't figure out the 3-way interaction to have models of
# Population x Year x Size, I'm splitting everything up by population
 BEstab_Plants <- Estab_Plants %>%
  filter(Population == "Bayfield")
FEstab_Plants <- Estab_Plants %>%
  filter(Population == "Fanshawe")
IEstab_Plants <- Estab_Plants %>%
  filter(Population == "Ilderton")
REstab_Plants <- Estab_Plants %>%
  filter(Population == "RBG")

BEstab_Plants2 <- Estab_Plants2 %>%
  filter(Population == "Bayfield")
FEstab_Plants2 <- Estab_Plants2 %>%
  filter(Population == "Fanshawe")
IEstab_Plants2 <- Estab_Plants2 %>%
  filter(Population == "Ilderton")
REstab_Plants2 <- Estab_Plants2 %>%
  filter(Population == "RBG")
```

Vital Rate Regressions and Probabilities

The code for these is mostly based on:
https://github.com/jloesberg/Primula/blob/main/scripts/primula_IPM.rmd

and also on Kuss et al. 2008 and Lemmer et al. 2021


Then run the regressions and set up parameters
```{r Regressions}
# Set bootstrap sample number
boot <- 1000

######## Probability of Germination ########################

#These numbers are from Emma Neigel's translocation experiment.

#Year 1 (one year after seeds fall--fresh seeds)
Prob_Germ_1 <- 105/2025 #about 5%

#Year 2 (two years after seeds fall--one year in seed bank)
Prob_Germ_2 <- 3/847 # 0.35%

# Seed viability (after first year)
Prob_Surv_Seed <- 58/803

######### Probability of Establishment #####################
#Seedling to Adult
est.intercepts <- sum(SurvData$Survival_t1[SurvData$Recruit_t == 1], na.rm = T)/sum(SurvData$Recruit_t, na.rm = T)

######### Distribution of Recruit Sizes ####################

#size (mean and sd) of recruits in year t_1 if they are a seedling that year
Recr_mean <- sapply(split(SurvData$Height_cm_t1[SurvData$Recruit_t1 == 1],
                          SurvData$Population[SurvData$Recruit_t1 == 1]), mean, na.rm = T) 
Recr_sd <- sapply(split(SurvData$Height_cm_t1[SurvData$Recruit_t1 == 1],
                        SurvData$Population[SurvData$Recruit_t1 == 1]), sd, na.rm = T) 

Recr_mean <- as.data.frame(Recr_mean)
Recr_mean$Population <- c(1:4)

Recr_sd <- as.data.frame(Recr_sd)
Recr_sd$Population <- c(1:4)

######### Adult Survival ###################################
# No significant interactions

# Run model
Surv_Model <- glm(Survival_t1 ~ Height_cm_t,
                  family = binomial(link = "logit"),
                  data = BEstab_Plants)
# Then get bootstrapped coefficients
set.seed(1)
Bs.samp <- MASS::mvrnorm(n = boot, mu = coef(Surv_Model), Sigma = vcov(Surv_Model))
# And also get model coefficients for deterministic model--later on
Bs.intercept <- Surv_Model$coefficients[1]
Bs.slope1 <- Surv_Model$coefficients[2]

# For Bayfield, the model doesn't converge, so I will not bootstrap survival
Bs.samp <- cbind(rep(Bs.intercept, 1000), rep(Bs.slope1, 1000))

Surv_Model <- glm(Survival_t1 ~ Height_cm_t,
                  family = binomial(link = "logit"),
                  data = FEstab_Plants)
set.seed(1)
Fs.samp <- MASS::mvrnorm(n = boot, mu = coef(Surv_Model), Sigma = vcov(Surv_Model))
Fs.intercept <- Surv_Model$coefficients[1]
Fs.slope1 <- Surv_Model$coefficients[2]

Surv_Model <- glm(Survival_t1 ~ Height_cm_t,
                  family = binomial(link = "logit"),
                  data = IEstab_Plants)
set.seed(1)
Is.samp <- MASS::mvrnorm(n = boot, mu = coef(Surv_Model), Sigma = vcov(Surv_Model))
Is.intercept <- Surv_Model$coefficients[1]
Is.slope1 <- Surv_Model$coefficients[2]

Surv_Model <- glm(Survival_t1 ~ Height_cm_t,
                  family = binomial(link = "logit"),
                  data = REstab_Plants)
set.seed(1)
Rs.samp <- MASS::mvrnorm(n = boot, mu = coef(Surv_Model), Sigma = vcov(Surv_Model))
Rs.intercept <- Surv_Model$coefficients[1]
Rs.slope1 <- Surv_Model$coefficients[2]

######### Adult Growth #####################################

Growth_Model <- lm(Height_cm_t1 ~ Year_t * Height_cm_t,
              data = BEstab_Plants)
set.seed(1)
Bg.samp <- MASS::mvrnorm(n = boot, mu = coef(Growth_Model), Sigma = vcov(Growth_Model))
Bg.intercept <- Growth_Model$coefficients[1]
Bg.slope1 <- Growth_Model$coefficients[2]
Bg.slope2 <- Growth_Model$coefficients[3]
Bg.slope3 <- Growth_Model$coefficients[4]
# variance
Bgrowth_var <- sd(resid(Growth_Model))

Growth_Model <- lm(Height_cm_t1 ~ Year_t * Height_cm_t,
              data = FEstab_Plants)
set.seed(1)
Fg.samp <- MASS::mvrnorm(n = boot, mu = coef(Growth_Model), Sigma = vcov(Growth_Model))
Fg.intercept <- Growth_Model$coefficients[1]
Fg.slope1 <- Growth_Model$coefficients[2]
Fg.slope2 <- Growth_Model$coefficients[3]
Fg.slope3 <- Growth_Model$coefficients[4]
# variance
Fgrowth_var <- sd(resid(Growth_Model))

Growth_Model <- lm(Height_cm_t1 ~ Year_t * Height_cm_t,
              data = IEstab_Plants)
set.seed(1)
Ig.samp <- MASS::mvrnorm(n = boot, mu = coef(Growth_Model), Sigma = vcov(Growth_Model))
Ig.intercept <- Growth_Model$coefficients[1]
Ig.slope1 <- Growth_Model$coefficients[2]
Ig.slope2 <- Growth_Model$coefficients[3]
Ig.slope3 <- Growth_Model$coefficients[4]
# variance
Igrowth_var <- sd(resid(Growth_Model))

Growth_Model <- lm(Height_cm_t1 ~ Year_t * Height_cm_t,
              data = REstab_Plants)
set.seed(1)
Rg.samp <- MASS::mvrnorm(n = boot, mu = coef(Growth_Model), Sigma = vcov(Growth_Model))
Rg.intercept <- Growth_Model$coefficients[1]
Rg.slope1 <- Growth_Model$coefficients[2]
Rg.slope2 <- Growth_Model$coefficients[3]
Rg.slope3 <- Growth_Model$coefficients[4]
# variance
Rgrowth_var <- sd(resid(Growth_Model))

########## Probability of Flowering ########################

#The variable "Flowering" is for time t+1
Flower_Model <- glm(Flowering ~ Year_t * Height_cm_t,
              family = binomial(link = "logit"),
              data = BEstab_Plants)
set.seed(1)
Bf.samp <- MASS::mvrnorm(n = boot, mu = coef(Flower_Model), Sigma = vcov(Flower_Model))
Bf.intercept <- Flower_Model$coefficients[1]
Bf.slope1 <- Flower_Model$coefficients[2]
Bf.slope2 <- Flower_Model$coefficients[3]
Bf.slope3 <- Flower_Model$coefficients[4]

Flower_Model <- glm(Flowering ~ Year_t * Height_cm_t,
              family = binomial(link = "logit"),
              data = FEstab_Plants)
set.seed(1)
Ff.samp <- MASS::mvrnorm(n = boot, mu = coef(Flower_Model), Sigma = vcov(Flower_Model))
Ff.intercept <- Flower_Model$coefficients[1]
Ff.slope1 <- Flower_Model$coefficients[2]
Ff.slope2 <- Flower_Model$coefficients[3]
Ff.slope3 <- Flower_Model$coefficients[4]

Flower_Model <- glm(Flowering ~ Year_t * Height_cm_t,
              family = binomial(link = "logit"),
              data = IEstab_Plants)
set.seed(1)
If.samp <- MASS::mvrnorm(n = boot, mu = coef(Flower_Model), Sigma = vcov(Flower_Model))
If.intercept <- Flower_Model$coefficients[1]
If.slope1 <- Flower_Model$coefficients[2]
If.slope2 <- Flower_Model$coefficients[3]
If.slope3 <- Flower_Model$coefficients[4]

Flower_Model <- glm(Flowering ~ Year_t * Height_cm_t,
              family = binomial(link = "logit"),
              data = REstab_Plants)
set.seed(1)
Rf.samp <- MASS::mvrnorm(n = boot, mu = coef(Flower_Model), Sigma = vcov(Flower_Model))
Rf.intercept <- Flower_Model$coefficients[1]
Rf.slope1 <- Flower_Model$coefficients[2]
Rf.slope2 <- Flower_Model$coefficients[3]
Rf.slope3 <- Flower_Model$coefficients[4]

######### Seeds Per Plant ##################################

#t1_x is the log of seed number in t1

Seed_Model <- glm(Seed_Number_t1_x ~ Year_t * Height_cm_t,
                  data = BEstab_Plants2)
set.seed(1)
Bse.samp <- MASS::mvrnorm(n = boot, mu = coef(Seed_Model), Sigma = vcov(Seed_Model))
Bse.intercept <- Seed_Model$coefficients[1]
Bse.slope1 <- Seed_Model$coefficients[2]
Bse.slope2 <- Seed_Model$coefficients[3]
Bse.slope3 <- Seed_Model$coefficients[4]

Seed_Model <- glm(Seed_Number_t1_x ~ Year_t * Height_cm_t,
                  data = FEstab_Plants2)
set.seed(1)
Fse.samp <- MASS::mvrnorm(n = boot, mu = coef(Seed_Model), Sigma = vcov(Seed_Model))
Fse.intercept <- Seed_Model$coefficients[1]
Fse.slope1 <- Seed_Model$coefficients[2]
Fse.slope2 <- Seed_Model$coefficients[3]
Fse.slope3 <- Seed_Model$coefficients[4]

Seed_Model <- glm(Seed_Number_t1_x ~ Year_t * Height_cm_t,
                  data = IEstab_Plants2)
set.seed(1)
Ise.samp <- MASS::mvrnorm(n = boot, mu = coef(Seed_Model), Sigma = vcov(Seed_Model))
Ise.intercept <- Seed_Model$coefficients[1]
Ise.slope1 <- Seed_Model$coefficients[2]
Ise.slope2 <- Seed_Model$coefficients[3]
Ise.slope3 <- Seed_Model$coefficients[4]

Seed_Model <- glm(Seed_Number_t1_x ~ Year_t * Height_cm_t,
                  data = REstab_Plants2)
set.seed(1)
Rse.samp <- MASS::mvrnorm(n = boot, mu = coef(Seed_Model), Sigma = vcov(Seed_Model))
Rse.intercept <- Seed_Model$coefficients[1]
Rse.slope1 <- Seed_Model$coefficients[2]
Rse.slope2 <- Seed_Model$coefficients[3]
Rse.slope3 <- Seed_Model$coefficients[4]
```

Extract coefficients for table
```{r}
# Re-run models above with separate names
## I have to re-run these so I don't mess up anything with the bootstrap
## below, So these models are ONLY for making a table of coefficients

Surv_ModelB <- glm(Survival_t1 ~ Height_cm_t,
                  family = binomial(link = "logit"),
                  data = BEstab_Plants)
Surv_ModelF <- glm(Survival_t1 ~ Height_cm_t,
                  family = binomial(link = "logit"),
                  data = FEstab_Plants)
Surv_ModelI <- glm(Survival_t1 ~ Height_cm_t,
                  family = binomial(link = "logit"),
                  data = IEstab_Plants)
Surv_ModelR <- glm(Survival_t1 ~ Height_cm_t,
                  family = binomial(link = "logit"),
                  data = REstab_Plants)

Grow_ModelB <- glm(Height_cm_t1 ~ Year_t * Height_cm_t,
              data = BEstab_Plants)
Grow_ModelF <- glm(Height_cm_t1 ~ Year_t * Height_cm_t,
              data = FEstab_Plants)
Grow_ModelI <- glm(Height_cm_t1 ~ Year_t * Height_cm_t,
              data = IEstab_Plants)
Grow_ModelR <- glm(Height_cm_t1 ~ Year_t * Height_cm_t,
              data = REstab_Plants)

Flower_ModelB <- glm(Flowering ~ Year_t * Height_cm_t,
              family = binomial(link = "logit"),
              data = BEstab_Plants)
Flower_ModelF <- glm(Flowering ~ Year_t * Height_cm_t,
              family = binomial(link = "logit"),
              data = FEstab_Plants)
Flower_ModelI <- glm(Flowering ~ Year_t * Height_cm_t,
              family = binomial(link = "logit"),
              data = IEstab_Plants)

Flower_ModelR <- glm(Flowering ~ Year_t * Height_cm_t,
              family = binomial(link = "logit"),
              data = REstab_Plants)

Seed_ModelB <- glm(Seed_Number_t1_x ~ Year_t * Height_cm_t,
                  data = BEstab_Plants2)
Seed_ModelF <- glm(Seed_Number_t1_x ~ Year_t * Height_cm_t,
                  data = FEstab_Plants2)
Seed_ModelI <- glm(Seed_Number_t1_x ~ Year_t * Height_cm_t,
                  data = IEstab_Plants2)
Seed_ModelR <- glm(Seed_Number_t1_x ~ Year_t * Height_cm_t,
                  data = REstab_Plants2)

### Gather coefficients

Coefs <- c(coef(Surv_ModelB), coef(Surv_ModelF), coef(Surv_ModelI), coef(Surv_ModelR),
           coef(Grow_ModelB), coef(Grow_ModelF),  coef(Grow_ModelI),  coef(Grow_ModelR),
           coef(Flower_ModelB), coef(Flower_ModelF), coef(Flower_ModelI), coef(Flower_ModelR),
           coef(Seed_ModelB), coef(Seed_ModelF), coef(Seed_ModelI), coef(Seed_ModelR))
SECoefs <- c(se.coef(Surv_ModelB), se.coef(Surv_ModelF), se.coef(Surv_ModelI), se.coef(Surv_ModelR),
           se.coef(Grow_ModelB), se.coef(Grow_ModelF),  se.coef(Grow_ModelI),  se.coef(Grow_ModelR),
           se.coef(Flower_ModelB), se.coef(Flower_ModelF), se.coef(Flower_ModelI), se.coef(Flower_ModelR),
           se.coef(Seed_ModelB), se.coef(Seed_ModelF), se.coef(Seed_ModelI), se.coef(Seed_ModelR))

DF <- c(NA, Anova(Surv_ModelB)$Df, NA, Anova(Surv_ModelF)$Df,NA, Anova(Surv_ModelI)$Df,NA, Anova(Surv_ModelR)$Df,
        NA, Anova(Grow_ModelB)$Df, NA, Anova(Grow_ModelF)$Df, NA, Anova(Grow_ModelI)$Df, NA, Anova(Grow_ModelR)$Df,
        NA, Anova(Flower_ModelB)$Df, NA, Anova(Flower_ModelF)$Df, NA, Anova(Flower_ModelI)$Df, NA, Anova(Flower_ModelR)$Df,
        NA, Anova(Seed_ModelB)$Df, NA, Anova(Seed_ModelF)$Df, NA, Anova(Seed_ModelI)$Df, NA, Anova(Seed_ModelR)$Df)

# Need to re-run if change order of things in GLMs above!!!
AICModel <- c(drop1(Surv_ModelB)$AIC, drop1(Surv_ModelF)$AIC, drop1(Surv_ModelI)$AIC, drop1(Surv_ModelR)$AIC,
              NA, NA, drop1(Grow_ModelB)$AIC, NA, NA, drop1(Grow_ModelF)$AIC, NA, NA, drop1(Grow_ModelI)$AIC, NA, NA, drop1(Grow_ModelR)$AIC,
              NA, NA, drop1(Flower_ModelB)$AIC, NA, NA, drop1(Flower_ModelF)$AIC, NA, NA, drop1(Flower_ModelI)$AIC, NA, NA, drop1(Flower_ModelR)$AIC,
              NA, NA, drop1(Seed_ModelB)$AIC, NA, NA, drop1(Seed_ModelF)$AIC, NA, NA, drop1(Seed_ModelI)$AIC, NA, NA, drop1(Seed_ModelR)$AIC)

LRT <- c(NA, Anova(Surv_ModelB)$`LR Chisq`,NA, Anova(Surv_ModelF)$`LR Chisq`, NA, Anova(Surv_ModelI)$`LR Chisq`, NA, Anova(Surv_ModelR)$`LR Chisq`,
         NA, Anova(Grow_ModelB)$`LR Chisq`, NA, Anova(Grow_ModelF)$`LR Chisq`, NA, Anova(Grow_ModelI)$`LR Chisq`, NA, Anova(Grow_ModelR)$`LR Chisq`,
         NA, Anova(Flower_ModelB)$`LR Chisq`, NA, Anova(Flower_ModelF)$`LR Chisq`, NA, Anova(Flower_ModelI)$`LR Chisq`, NA, Anova(Flower_ModelR)$`LR Chisq`,
         NA, Anova(Seed_ModelB)$`LR Chisq`, NA, Anova(Seed_ModelF)$`LR Chisq`, NA, Anova(Seed_ModelI)$`LR Chisq`, NA, Anova(Seed_ModelR)$`LR Chisq`)

PValues <- c(NA, Anova(Surv_ModelB)$`Pr(>Chisq)`,NA, Anova(Surv_ModelF)$`Pr(>Chisq)`, NA, Anova(Surv_ModelI)$`Pr(>Chisq)`, NA, Anova(Surv_ModelR)$`Pr(>Chisq)`,
             NA, Anova(Grow_ModelB)$`Pr(>Chisq)`, NA, Anova(Grow_ModelF)$`Pr(>Chisq)`, NA, Anova(Grow_ModelI)$`Pr(>Chisq)`, NA, Anova(Grow_ModelR)$`Pr(>Chisq)`,
             NA, Anova(Flower_ModelB)$`Pr(>Chisq)`, NA, Anova(Flower_ModelF)$`Pr(>Chisq)`, NA, Anova(Flower_ModelI)$`Pr(>Chisq)`, NA, Anova(Flower_ModelR)$`Pr(>Chisq)`,
             NA, Anova(Seed_ModelB)$`Pr(>Chisq)`, NA, Anova(Seed_ModelF)$`Pr(>Chisq)`, NA, Anova(Seed_ModelI)$`Pr(>Chisq)`, NA, Anova(Seed_ModelR)$`Pr(>Chisq)`)

AllCoefs <- as.data.frame(cbind(Coefs, SECoefs, DF, AICModel, LRT, PValues))
names(AllCoefs) <- c("Coefficient", "SE", "df", "AIC", "Test_Statistic", "P_value")
AllCoefs$Predictor <- rownames(AllCoefs)
AllCoefs <- AllCoefs %>%
  dplyr::select(Predictor, Coefficient, SE, df, AIC, Test_Statistic, P_value) %>%
  mutate(Predictor = case_when(grepl("Intercept", Predictor) ~ "Intercept",
                               grepl("Year_t.Height", Predictor) ~ "Height Year t * Year t",
                               grepl("Year_t.", Predictor) ~ "Year t",
                               Predictor == "Year_t" ~ "Year_t",
                               TRUE ~ "Height Year t")) %>%
  mutate(across(c("Coefficient", "SE", "AIC", "Test_Statistic"), round, 2)) %>%
  mutate(across(c("P_value"), round, 3)) %>%
  add_row(Predictor = "Survival Model", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 1) %>%
  add_row(Predictor = "Bayfield", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 2) %>%
  add_row(Predictor = "Fanshawe", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 5) %>%
  add_row(Predictor = "Ilderton", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 8) %>%
  add_row(Predictor = "RBG", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 11) %>%
  
  add_row(Predictor = "Growth Model", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 14) %>%
  add_row(Predictor = "Bayfield", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 15) %>%
  add_row(Predictor = "Fanshawe", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 20) %>%
  add_row(Predictor = "Ilderton", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 25) %>%
  add_row(Predictor = "RBG", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 30) %>%
  
  add_row(Predictor = "Flowering Model", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 35) %>%
  add_row(Predictor = "Bayfield", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 36) %>%
  add_row(Predictor = "Fanshawe", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 41) %>%
  add_row(Predictor = "Ilderton", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 46) %>%
  add_row(Predictor = "RBG", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 51) %>%
  
  add_row(Predictor = "Seed Model", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 56) %>%
  add_row(Predictor = "Bayfield", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 57) %>%
  add_row(Predictor = "Fanshawe", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 62) %>%
  add_row(Predictor = "Ilderton", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 67) %>%
  add_row(Predictor = "RBG", Coefficient = NA, SE = NA, df = NA, AIC = NA, Test_Statistic = NA, P_value = NA, .before = 72)
# write.csv(AllCoefs, "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs/PopCoefs.csv", row.names = F)
```

# Graphs of Vital Rates

Plot vital rates (probability of survival, growth, probability of
 flowering, and seeds per plant)
```{r echo=FALSE}
#Make a vector of population names
Population <- c("Bayfield", "Fanshawe", "Ilderton", "RBG")
PopCol <- c("#D55E00", "#0072B2", "#F0E442","#009E73")
pch_names <- c(1:4)

## Growth
p1 <- ggplot(Estab_Plants, aes(x = Height_cm_t, y = Height_cm_t1,
                         col = Population, linetype = Population)) +
  geom_point(alpha = 0.5) +
  labs (x = "Height in year t (cm)", y = "Height in year t + 1 (cm)") +
  scale_color_manual(values = PopCol) +
  stat_smooth(method = "lm", formula = y ~ x, size = 1, se = T, fill = NA) +
    scale_linetype_manual(values = c("Bayfield" = 1, "Fanshawe" = 1, "Ilderton" = 1, "RBG" = 1)) +
  xlim(0, 90) +
  theme_classic() +
  theme(legend.position="none") +
    theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  # Fix font sizes
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14))

## Flowering

#divide by population
Bayfield <- Estab_Plants %>%
  filter(Population == "Bayfield")
Fanshawe <- Estab_Plants %>%
  filter(Population == "Fanshawe")
Ilderton <- Estab_Plants %>%
  filter(Population == "Ilderton")
RBG <- Estab_Plants %>%
  filter(Population == "RBG")

#create dataframe of points along regression for ggplot
fitB = glm(Flowering ~ Height_cm_t, data = Bayfield, family=binomial)
newdatB <- data.frame(Height_cm_t =seq(min(Estab_Plants$Height_cm_t, na.rm = T)*0.9,
                                       max(Estab_Plants$Height_cm_t, na.rm = T)*1.1,len=1000))
newdatB$Flowering = predict(fitB, newdata=newdatB, type="response")

fitF = glm(Flowering ~ Height_cm_t, data = Fanshawe, family=binomial)
newdatF <- data.frame(Height_cm_t =seq(min(Estab_Plants$Height_cm_t, na.rm = T)*0.9,
                                       max(Estab_Plants$Height_cm_t, na.rm = T)*1.1,len=1000))
newdatF$Flowering = predict(fitF, newdata=newdatF, type="response")

fitI = glm(Flowering ~ Height_cm_t, data = Ilderton, family=binomial)
newdatI <- data.frame(Height_cm_t =seq(min(Estab_Plants$Height_cm_t, na.rm = T)*0.9,
                                       max(Estab_Plants$Height_cm_t, na.rm = T)*1.1,len=1000))
newdatI$Flowering = predict(fitI, newdata=newdatI, type="response")

fitR = glm(Flowering ~ Height_cm_t, data = RBG, family=binomial)
newdatR <- data.frame(Height_cm_t =seq(min(Estab_Plants$Height_cm_t, na.rm = T)*0.9,
                                       max(Estab_Plants$Height_cm_t, na.rm = T)*1.1,len=1000))
newdatR$Flowering = predict(fitR, newdata=newdatR, type="response")

#plot!
p2 <- ggplot(Estab_Plants, aes(x = Height_cm_t, y = Flowering, col = Population)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = PopCol) +
  geom_line(data = newdatB, aes(x = Height_cm_t, y = Flowering), col = "#D55E00", linetype = 1, linewidth = 1) +
  geom_line(data = newdatF, aes(x = Height_cm_t, y = Flowering), col = "#0072B2", linetype = 1, linewidth = 1) +
  geom_line(data = newdatI, aes(x = Height_cm_t, y = Flowering), col = "#F0E442", linetype = 1, linewidth = 1) +
  geom_line(data = newdatR, aes(x = Height_cm_t, y = Flowering), col = "#009E73", linetype = 1, linewidth = 1) +
  labs(x = "Height in year t (cm)", y = "Probability of Flowering Year t+1") +
  xlim(0, 90) +
  theme_classic() + 
  theme(legend.position="none") +
  # Fix font sizes
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14))

# Survival

#divide by population
Bayfield <- Estab_Plants %>%
  filter(Population == "Bayfield")
Fanshawe <- Estab_Plants %>%
  filter(Population == "Fanshawe")
Ilderton <- Estab_Plants %>%
  filter(Population == "Ilderton")
RBG <- Estab_Plants %>%
  filter(Population == "RBG")

#create dataframe of points along regression for ggplot
fitB = glm(Survival_t1 ~ Height_cm_t, data = Bayfield, family=binomial)
newdatB <- data.frame(Height_cm_t =seq(min(Estab_Plants$Height_cm_t, na.rm = T)*0.9,
                                       max(Estab_Plants$Height_cm_t, na.rm = T)*1.1,len=1000))
newdatB$Survival_t1 = predict(fitB, newdata=newdatB, type="response")

fitF = glm(Survival_t1 ~ Height_cm_t, data = Fanshawe, family=binomial)
newdatF <- data.frame(Height_cm_t =seq(min(Estab_Plants$Height_cm_t, na.rm = T)*0.9,
                                       max(Estab_Plants$Height_cm_t, na.rm = T)*1.1,len=1000))
newdatF$Survival_t1 = predict(fitF, newdata=newdatF, type="response")

fitI = glm(Survival_t1 ~ Height_cm_t, data = Ilderton, family=binomial)
newdatI <- data.frame(Height_cm_t =seq(min(Estab_Plants$Height_cm_t, na.rm = T)*0.9,
                                       max(Estab_Plants$Height_cm_t, na.rm = T)*1.1,len=1000))
newdatI$Survival_t1 = predict(fitI, newdata=newdatI, type="response")

fitR = glm(Survival_t1 ~ Height_cm_t, data = RBG, family=binomial)
newdatR <- data.frame(Height_cm_t =seq(min(Estab_Plants$Height_cm_t, na.rm = T)*0.9,
                                       max(Estab_Plants$Height_cm_t, na.rm = T)*1.1,len=1000))
newdatR$Survival_t1 = predict(fitR, newdata=newdatR, type="response")

#merge back
newdatB$Population <- "Bayfield"
newdatF$Population <- "Fanshawe"
newdatI$Population <- "Ilderton"
newdatR$Population <- "RBG"
newdat <- as.data.frame(rbind(newdatB, newdatF, newdatI, newdatR))

#plot!
p4 <- ggplot(Estab_Plants, aes(x = Height_cm_t, y = Survival_t1, col = Population)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = PopCol, # labels used to change legend
                     #labels=c("1", "3", "4", "6")
                     labels = c("Bayfield", "Fanshawe", "Ilderton", "RBG")) +
  geom_line(data = newdat, aes(x = Height_cm_t, y = Survival_t1, linetype = Population), linewidth = 1) +
  labs(x = "Height in year t (cm)", y = "Probability of Survival Year t+1") +
  scale_linetype_manual(values = c("Bayfield" = 1, "Fanshawe" = 1, "Ilderton" = 1, "RBG" = 1),
                        #labels=c("1", "3", "4", "6")
                        labels = c("Bayfield", "Fanshawe", "Ilderton", "RBG")) +
  xlim(0, 90) +
  theme_classic() + 
  #theme(legend.position="none") +
  #add legend for all the graphs
  guides(colour = guide_legend(position = "inside")) +
  theme(legend.position.inside = c(0.85, 0.5),
        legend.background = element_rect(fill = "white", colour = "lightgray")) +
  theme(legend.key.size = unit(1, 'cm'),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))+
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  # Fix font sizes
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14))

# Seeds

p3 <- ggplot(Estab_Plants2, aes(x = Height_cm_t, y = log(Seed_Number_t1),
                         col = Population, linetype = Population)) +
  geom_point(alpha = 0.5) +
  labs (x = "Height in year t (cm)", y = "log(Seeds per plant), Year t + 1") +
  scale_color_manual(values = PopCol) +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1, se = T, fill = NA) +
  scale_linetype_manual(values = c("Bayfield" = 1, "Fanshawe" = 1, "Ilderton" = 1, "RBG" = 1)) +
  xlim(0, 90) +
  theme_classic() +
  theme(legend.position="none") +
  # Fix font sizes
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14))

Graphs <- ggarrange(p4, p1, p2, p3,
                    labels = c("(a)", "(b)", "(c)", "(d)"),
                    font.label = list(size = 12), 
                    hjust = -0.2,
                    ncol = 2, nrow = 2, align = "h")
Graphs

# ggsave(plot = Graphs, dpi = 600, width = 12, height = 8, # was 12w x 6 h
#        filename = "VitalRates_Pops.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = Graphs, dpi = 600, width = 12, height = 8, # was 12w x 6 h
#        filename = "VitalRates_Pops.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
```

# Bootstrapped Model

Once all the regressions have been run, combine the parameter estimates
 from them (slopes and intercepts, and variation), along with the
 probabilities of transition between stages.

Bootstrap code is from https://github.com/jloesberg/Primula/blob/main/scripts/primula_IPM.rmd
 
```{r Gather Parameters}
# First need year and population numbers
Year_t <- c(1,2,3)
Population <- c(1,2,3,4)

# Survival
# Combine the populations
s.samp <- rbind(Bs.samp, Fs.samp, Is.samp, Rs.samp)

# For slopes 1 and 2, just need 1 column of data
s.slope <- data.frame(s.slope = s.samp[, 2], Population = c(rep(1,1000), rep(2,1000),rep(3,1000),rep(4,1000)), boot.group = seq(1, boot, by = 1))

# Make an empty dataframe and set the row indexing
Bs.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Bs.int[row_index, "s.int"] <-  Bs.samp[j, 1]# + Bs.samp[j, 3] * Year_t[i]
    Bs.int[row_index, "Year_t"] <- Year_t[i]
    Bs.int[row_index, "Population"] <- 1
    Bs.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}
Fs.int <- data.frame()
row_index <- 1
for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Fs.int[row_index, "s.int"] <-  Fs.samp[j, 1]# + Fs.samp[j, 3] * Year_t[i]
    Fs.int[row_index, "Year_t"] <- Year_t[i]
    Fs.int[row_index, "Population"] <- 2
    Fs.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}
Is.int <- data.frame()
row_index <- 1
for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Is.int[row_index, "s.int"] <-  Is.samp[j, 1]# + Is.samp[j, 3] * Year_t[i]
    Is.int[row_index, "Year_t"] <- Year_t[i]
    Is.int[row_index, "Population"] <- 3
    Is.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}
Rs.int <- data.frame()
row_index <- 1
for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Rs.int[row_index, "s.int"] <-  Rs.samp[j, 1]# + Rs.samp[j, 3] * Year_t[i]
    Rs.int[row_index, "Year_t"] <- Year_t[i]
    Rs.int[row_index, "Population"] <- 4
    Rs.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

s.int <- as.data.frame(rbind(Bs.int, Fs.int, Is.int, Rs.int))

surv <-left_join(s.int, s.slope, by = c("Population", "boot.group"))
# Growth

# Combine the populations
g.samp <- rbind(Bg.samp, Fg.samp, Ig.samp, Rg.samp)

# For slopes 1 and 2, just need 1 column of data
g.slope <- data.frame(g.slope = g.samp[, 3], Population = c(rep(1,1000), rep(2,1000),rep(3,1000),rep(4,1000)), boot.group = seq(1, boot, by = 1))
g.slope2 <- data.frame(g.slope2 = g.samp[, 4], Population = c(rep(1,1000), rep(2,1000),rep(3,1000),rep(4,1000)), boot.group = seq(1, boot, by = 1))

# Make an empty dataframe and set the row indexing
Bg.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Bg.int[row_index, "g.int"] <-  Bg.samp[j, 1] + Bg.samp[j, 2] * Year_t[i]
    Bg.int[row_index, "Year_t"] <- Year_t[i]
    Bg.int[row_index, "Population"] <- 1
    Bg.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

Fg.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Fg.int[row_index, "g.int"] <-  Fg.samp[j, 1] + Fg.samp[j, 2] * Year_t[i]
    Fg.int[row_index, "Year_t"] <- Year_t[i]
    Fg.int[row_index, "Population"] <- 2
    Fg.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
    }
  }

Ig.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Ig.int[row_index, "g.int"] <-  Ig.samp[j, 1] + Ig.samp[j, 2] * Year_t[i]
    Ig.int[row_index, "Year_t"] <- Year_t[i]
    Ig.int[row_index, "Population"] <- 3
    Ig.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
    }
}

Rg.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Rg.int[row_index, "g.int"] <-  Rg.samp[j, 1] + Rg.samp[j, 2] * Year_t[i]
    Rg.int[row_index, "Year_t"] <- Year_t[i]
    Rg.int[row_index, "Population"] <- 4
    Rg.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

g.int <- as.data.frame(rbind(Bg.int, Fg.int, Ig.int, Rg.int))

# Combine the slopes and intercept
grow <-left_join(g.int, g.slope, by = c("Population", "boot.group"))
grow <-left_join(grow, g.slope2, by = c("Population", "boot.group"))

# Growth Variance
# For the growth parameter, need variation
growth_var <- data.frame(Grow_sd = double(), Population = double())
growth_var[1,1] <- Bgrowth_var
growth_var[2,1] <- Fgrowth_var
growth_var[3,1] <- Igrowth_var
growth_var[4,1] <- Rgrowth_var
growth_var[1,2] <- 1 # Bayfield
growth_var[2,2] <- 2 # Fanshawe
growth_var[3,2] <- 3 # Ilderton
growth_var[4,2] <- 4 # RBG

# Flowers
f.samp <- rbind(Bf.samp, Ff.samp, If.samp, Rf.samp)

f.slope <- data.frame(f.slope = f.samp[, 3], Population = c(rep(1,1000), rep(2,1000),rep(3,1000),rep(4,1000)), boot.group = seq(1, boot, by = 1))
f.slope2 <- data.frame(f.slope2 = f.samp[, 4], Population = c(rep(1,1000), rep(2,1000),rep(3,1000),rep(4,1000)), boot.group = seq(1, boot, by = 1))

Bf.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Bf.int[row_index, "fl.int"] <-  Bf.samp[j, 1] + Bf.samp[j, 2] * Year_t[i]
    Bf.int[row_index, "Year_t"] <- Year_t[i]
    Bf.int[row_index, "Population"] <- 1
    Bf.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}
Ff.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Ff.int[row_index, "fl.int"] <-  Ff.samp[j, 1] + Ff.samp[j, 2] * Year_t[i]
    Ff.int[row_index, "Year_t"] <- Year_t[i]
    Ff.int[row_index, "Population"] <- 2
    Ff.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

If.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    If.int[row_index, "fl.int"] <-  If.samp[j, 1] + If.samp[j, 2] * Year_t[i]
    If.int[row_index, "Year_t"] <- Year_t[i]
    If.int[row_index, "Population"] <- 3
    If.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

Rf.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Rf.int[row_index, "fl.int"] <-  Rf.samp[j, 1] + Rf.samp[j, 2] * Year_t[i]
    Rf.int[row_index, "Year_t"] <- Year_t[i]
    Rf.int[row_index, "Population"] <- 4
    Rf.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

f.int <- as.data.frame(rbind(Bf.int, Ff.int, If.int, Rf.int))

fl<-left_join(f.int, f.slope, by = c("Population", "boot.group"))
fl <-left_join(fl, f.slope2, by = c("Population", "boot.group"))

# Seeds
se.samp <- rbind(Bse.samp, Fse.samp, Ise.samp, Rse.samp)

se.slope <- data.frame(se.slope = se.samp[, 3], Population = c(rep(1,1000), rep(2,1000),rep(3,1000),rep(4,1000)), boot.group = seq(1, boot, by = 1))
se.slope2 <- data.frame(se.slope2 = se.samp[, 4], Population = c(rep(1,1000), rep(2,1000),rep(3,1000),rep(4,1000)), boot.group = seq(1, boot, by = 1))

Bse.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Bse.int[row_index, "se.int"] <-  Bse.samp[j, 1] + Bse.samp[j, 2] * Year_t[i]
    Bse.int[row_index, "Year_t"] <- Year_t[i]
    Bse.int[row_index, "Population"] <- 1
    Bse.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

Fse.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Fse.int[row_index, "se.int"] <-  Fse.samp[j, 1] + Fse.samp[j, 2] * Year_t[i]
    Fse.int[row_index, "Year_t"] <- Year_t[i]
    Fse.int[row_index, "Population"] <- 2
    Fse.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

Ise.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Ise.int[row_index, "se.int"] <-  Ise.samp[j, 1] + Ise.samp[j, 2] * Year_t[i]
    Ise.int[row_index, "Year_t"] <- Year_t[i]
    Ise.int[row_index, "Population"] <- 3
    Ise.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

Rse.int <- data.frame()
row_index <- 1

for (j in 1:boot) {
  for (i in 1:length(Year_t)) {
    Rse.int[row_index, "se.int"] <-  Rse.samp[j, 1] + Rse.samp[j, 2] * Year_t[i]
    Rse.int[row_index, "Year_t"] <- Year_t[i]
    Rse.int[row_index, "Population"] <- 4
    Rse.int[row_index, "boot.group"] <- j
    # move to the next row
    row_index <- row_index + 1
  }
}

se.int <- as.data.frame(rbind(Bse.int, Fse.int, Ise.int, Rse.int))

se <-left_join(se.int, se.slope, by = c("Population", "boot.group"))
se <-left_join(se, se.slope2, by = c("Population", "boot.group"))

### Combine them all into one dataframe
boot_params <- left_join(surv, grow)
boot_params <- left_join(boot_params, fl, by = c("Year_t", "Population", "boot.group"))
boot_params <- left_join(boot_params, se, by = c("Year_t", "Population", "boot.group"))

### Add other things that aren't bootstrapped--these are considered to be the
# same across all populations and years
boot_params$Prob_Germ_1 <- Prob_Germ_1
boot_params$Prob_Germ_2 <- Prob_Germ_2
boot_params$Prob_Surv_Seed <- Prob_Surv_Seed
boot_params$Estab <- est.intercepts
boot_params$Recruit_size <- Recr_mean$Recr_mean[match(boot_params$Population, Recr_mean$Population)]
boot_params$Recruit_sd <- Recr_sd$Recr_sd[match(boot_params$Population, Recr_sd$Population)]

# Add the growth variation--this is considered to be the same across years
#  but can be different between the populations
boot_params$grow.sd <- growth_var$Grow_sd[match(boot_params$Population, growth_var$Population)]

# Make into a list, one element for each row
params_list <-  filter(boot_params) %>%
  split(., seq(nrow(.)))
```

## Functions

This is from a mix of Merow et al. 2014, the Merow github page, Kuss et al.
 2008 and Lemmer et al. 2021
 
Useful code pages:
https://github.com/Martin19910130/Bromus_IPM_publication/blob/master/R_code/Bootstrap_lambda.R
https://cmerow.github.io/RDataScience/21_Intro_IPMs.html#3_regressions_for_vital_rate_functions

```{r Functions}
########## Eviction ########################################
# Avoid unintentional eviction
# Code for this is from Lemmer et al. 2021
#Basically: force anything bigger than the max.size to be 
# the max.size, and anything smaller than the min.size to
# be the min.size
#Set minimum and maximum sizes
min.size = 0.9 * min(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
max.size = 1.1 * max(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)

x_range <- function(Height_cm_t, boot_params) {
  pmin(pmax(Height_cm_t, min.size), max.size)
}

######### Survival #########################################
#Survival of established plants
S_z <- function(Height_cm_t, Year_t, boot_params) {
  xb <- x_range(Height_cm_t, boot_params)
  linear_p <- boot_params$s.int + boot_params$s.slope * xb
  p <- 1/(1 + exp(-linear_p))
  return(p)
}

#Test this
mean(S_z(12, 1, boot_params))
mean(S_z(70, 2, boot_params))
#Seems ok

######### Growth ###########################################

G_z1z <- function(Height_cm_t, Height_cm_t1, Year_t, boot_params) {
  xb <- x_range(Height_cm_t, Params)
  return(dnorm(Height_cm_t1,
  mean <- boot_params$g.int + boot_params$g.slope * xb + boot_params$g.slope2 * Year_t * xb,
  sd = boot_params$grow.sd))
}

mean(G_z1z(12, 15, 1, boot_params))
mean(G_z1z(62, 70, 2, boot_params))
# I think this is right

######### Combine Survival and Growth ######################
P_z <- function(Height_cm_t, Height_cm_t1, Year_t, boot_params){
  xb <- x_range(Height_cm_t, boot_params)
  return(S_z(xb, Year_t, boot_params) * G_z1z(xb, Height_cm_t1, Year_t, boot_params))
}

mean(P_z(12, 15, 1, boot_params))
mean(P_z(54, 72, 2, boot_params))

######### Probability of Flowering #########################

Pf_z <- function(Height_cm_t, Year_t, boot_params) {
  xb <- x_range(Height_cm_t, boot_params)
  u <- boot_params$fl.int + boot_params$f.slope * xb + boot_params$f.slope2 * Year_t * xb
  p <- 1/(1 + exp(-u))
  return(p)
}

mean(Pf_z(12, 1, boot_params))
mean(Pf_z(72, 2, boot_params))
# Yup

######### Seeds Per Plant ##################################

Sf_z <- function(Height_cm_t, Year_t, boot_params){
  nseed <- boot_params$se.int + boot_params$se.slope * Height_cm_t + boot_params$se.slope2 * Year_t * Height_cm_t
  return(exp(nseed))
}

# Test this
mean(Sf_z(12, 1, boot_params))
mean(Sf_z(72, 2, boot_params))
# Yup!!

######### Flowering Probability and Seed Number ############

F_z <- function(Height_cm_t, Year_t, boot_params) {
  xb <- x_range(Height_cm_t, boot_params) # from Lemmer et al. 2021
  x <- Pf_z(xb, Year_t, boot_params) * Sf_z(xb, Year_t, boot_params)
  return(x)
}

# Test this
mean(F_z(12, 1, boot_params))
mean(F_z(72, 2, boot_params))

########## Seedling Sizes ##################################
Se_z <- function(Height_cm_t1, boot_params){
  dnorm(x = Height_cm_t1,
        mean = boot_params$Recruit_size,
        sd = boot_params$Recruit_sd)
}

#Test this
mean(Se_z(5, boot_params))
mean(Se_z(10, boot_params))
# Yeah

########## Seedling Survival ###############################
Sur_z1 <- function(boot_params){
  surv_adults <- boot_params$Estab
  return(surv_adults)
}

#Test this
mean(Sur_z1(boot_params))
#yep

########## Seed bank #######################################
S1_z <- function(Height_cm_t, Year_t, boot_params){
  f <- F_z(Height_cm_t, Year_t, boot_params) * (1 - boot_params$Prob_Germ_1)
  return(f);
}

#Test this
mean(S1_z(12, 1, boot_params))
mean(S1_z(72, 2, boot_params))
#looks pretty good
```

## Kernel

This is building the matrix for analysis. The fertility matrix and the
 transition matrix each fill specific cells of the "big matrix." The
 matrix has 100 rows and columns for the continuous stage, and then one
 row and column each for the discrete stages.
 The matrix is built as a function so that each population can be
  analysed separately in the next step.
 
 Big Matrix
 
 Rows are year t + 1
 Columns are year t
 
Based on code from Jenna Loesberg
https://github.com/jloesberg/Primula/blob/main/scripts/primula_IPM.rmd

One issue here: what does the transpose of the outer() function do? It really
 changes the growth rates, but some code I've found online and in textbooks and
 papers uses it, while others do exactly the same thing but without the transpose.
 I have decided not to use it, following Merow et al. 2014 Appendix A.

```{r Kernel, warning=FALSE}
#Set minimum and maximum sizes
min.size <- 0.9 * min(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
max.size <- 1.1 * max(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)

#building off of code from Tom Miller:
mat.size = 100 #dimensions of continuous matrix

# Three-year seedbank
bigmatrix <- function(params_list, mat.size){
  n <- mat.size #matrix size
  L <- min.size #lower integration limit
  U <- max.size #upper integration limit
  
  h <- (U - L)/n #Bin size
  b <- L + c(0:n) * h #Lower boundaries 
  y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
  # y is the mesh points of the matrix
  
  # Fertility matrix
  Fmat<-matrix(0,(n+4),(n+4))
  
  # Fecundity
  
  #Seedlings this year
  Fmat[4,5:(n+4)] <- Prob_Germ_1 * F_z(y, Year_t, params_list) + #seedlings this years seed
    Prob_Germ_2 * Prob_Surv_Seed * S1_z(y, Year_t, params_list) + #last year seed
    Prob_Germ_2 * (1 - Prob_Germ_1) * Prob_Surv_Seed * Prob_Germ_2 * Prob_Surv_Seed *  S1_z(y, Year_t, params_list) #2 year ago seed
  
  #Seedbank with t-1 and t-2 seeds
  Fmat[1,5:(n+4)] <- (1 - Prob_Germ_1) * (1 - Prob_Germ_2) * Prob_Surv_Seed * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #3-year old seeds
  Fmat[2,5:(n+4)] <- (1 - Prob_Germ_1) * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #2-year old seeds
  Fmat[3,5:(n+4)] <- (1 - Prob_Germ_1 - Prob_Germ_2) * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #1-year old seeds
  
  # Growth/survival transition matrix
  Tmat<-matrix(0,(n+4),(n+4))
  
  # probability a seed stays in the seedbank
  Tmat[2,1] <- (1 - Prob_Germ_1) * Prob_Surv_Seed
  Tmat[3,1] <- (1 - Prob_Germ_1) * Prob_Surv_Seed * (1 - Prob_Germ_2) * Prob_Surv_Seed
  Tmat[3,2] <- (1 - Prob_Germ_1) * Prob_Surv_Seed
  
  # probability seed germinates after 2 years
  Tmat[4,1] <- Prob_Germ_2 * Prob_Surv_Seed * (1 - Prob_Germ_2) * Prob_Surv_Seed
  Tmat[4,2] <- Prob_Germ_2 * Prob_Surv_Seed
  
  # probability seed germinates fresh
  Tmat[4,3] <- Prob_Germ_1
  
  # probability a seedling establishes as adult
  Tmat[5:(n+4),4] <- Sur_z1(params_list) * Se_z(y, params_list)
  
  # survival and growth
  Tmat[5:(n+4),5:(n+4)] <- (outer(y,y,P_z, Year_t, params_list)) * h #removed transpose of outer()
  
  # Full kernel -- adds two kernels built
  IPMmat <- Fmat + Tmat

  out_lamb <- data.frame(lambda = lambda(IPMmat))
  return(out_lamb)
}

Year_t <- c(1,2,3)

# Run the bigmatrix for each row of Params to get lambda
lamb_lists <- lapply(X = params_list, FUN = bigmatrix, mat.size = mat.size)
```

Check that everything looks right
This is a way to check if I needed the transpose function for the outer() function
I'm just running the bigmatrix once, randomly using the first row of boot_params
```{r Check output}
n <- mat.size #matrix size
L <- min.size #lower integration limit
U <- max.size #upper integration limit
h <- (U - L)/n #Bin size
b <- L + c(0:n) * h #Lower boundaries 
y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints

Fmat<-matrix(0,(n+3),(n+3))
# Fecundity
#Seedlings this year
Fmat[3,4:(n+3)] <- Prob_Germ_1 * F_z(y, Year_t, boot_params[1,]) + #seedlings this years seed
  Prob_Germ_2 * Prob_Surv_Seed * S1_z(y, Year_t, boot_params[1,]) + #last year seed
  Prob_Germ_2 * (1 - Prob_Germ_1) * Prob_Surv_Seed * Prob_Germ_2 * Prob_Surv_Seed *  S1_z(y, Year_t, boot_params[1,]) #2 year ago seed
#Seedbank with t-1 and t-2 seeds
Fmat[1,4:(n+3)] <- (1 - Prob_Germ_1) * Prob_Surv_Seed * S1_z(y, Year_t, boot_params[1,]) #2-year old seeds
Fmat[2,4:(n+3)] <- (1 - Prob_Germ_1 - Prob_Germ_2) * Prob_Surv_Seed * S1_z(y, Year_t, boot_params[1,]) #1-year old seeds
# Growth/survival transition matrix
Tmat<-matrix(0,(n+3),(n+3))
# probability a seed stays in the seedbank
Tmat[2,1] <- 1 - Prob_Germ_1 * Prob_Surv_Seed
# probability seed germinates after 2 years
Tmat[3,1] <- Prob_Germ_2 * Prob_Surv_Seed

# probability seed germinates fresh
Tmat[3,2] <- Prob_Germ_1
# probability a seedling establishes as adult
Tmat[4:(n+3),3] <- Sur_z1(boot_params[1,]) * Se_z(y, boot_params[1,])
# survival and growth
Tmat[4:(n+3),4:(n+3)] <- (outer(y,y,P_z, Year_t, boot_params[1,])) * h #removed transpose of outer()
# Full kernel -- adds two kernels built
IPMmat <- Fmat + (Tmat)

image(Tmat)

image((Tmat^.5))
image(t(Tmat^.5))

image((Fmat^.5))
image(t(Fmat^.5))

# Another check:

# The main issue is the order of year t and year t1 in the functions
# In mine:
G_z1z <- function(Height_cm_t, Height_cm_t1, Year_t, boot_params) {
  xb <- x_range(Height_cm_t, Params) # from Lemmer et al. 2021
  return(dnorm(Height_cm_t1,
  mean <- boot_params$g.int + boot_params$g.slope * xb + boot_params$g.slope2 * Year_t * xb,
  sd = boot_params$grow.sd))
}
# Year t comes first

# Based on the Appendix 1A page 12 in Merow et al. 2014 ("Advancing Population Ecology...),
# this is also correct
# They do the reverse (t1 and then t), but also transpose later on
```
I think it looks like I don't need the transpose for the way I've written the growth
 and survival function P_z. It makes more sense for a plant to get bigger from
 one year to the next than to get smaller or stay the same size, especially for
 the smaller plants.

Save lambda
```{r eval=FALSE, include=FALSE}
#save model output
# saveRDS(lamb_lists, file = "~/Research/Data/IPM_Formatted_STYL/Output/IPM_Lambda.rds")

# Load IPM output
# lamb_lists <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPM_Lambda.rds")
```

Summarize the bootstrapped growth rates
```{r}
boot_lambda <- do.call(rbind, lamb_lists)

boot_lambda[,2] <- c(rep(1, 3000), rep(2, 3000), rep(3, 3000), rep(4, 3000))
boot_lambda[,3] <- rep(c(1, 2, 3), times = 4000)

names(boot_lambda) <- c("lambda", "Population", "Year_t")
str(boot_lambda)
boot_lambda <- boot_lambda %>%
  mutate(Population = case_when(Population == 1 ~ "Bayfield",
                                Population == 2 ~ "Fanshawe",
                                Population == 3 ~ "Ilderton",
                                Population == 4 ~ "RBG"),
         Year_t = case_when(Year_t == 1 ~ 2022,
                            Year_t == 2 ~ 2023,
                            Year_t == 3 ~ 2024))
boot_lambda$Population <- as.factor(boot_lambda$Population)
boot_lambda$Year_t <- as.factor(boot_lambda$Year_t)

head(boot_lambda)

# Find mean and standard error for each population in each year
boot_error <- boot_lambda %>%
  group_by(Population, Year_t) %>%
  summarise(SE = sd(lambda)/sqrt(length(lambda)),
            Mean = mean(lambda)) %>%
  arrange(Mean)
boot_error
```

Plot growth rates
```{r}
# make facet labels
PopFacet <- c("Bayfield" = "(a)", "Fanshawe" = "(b)", "Ilderton"= "(c)", "RBG" = "(d)")

# As a boxplot
p <- ggplot(boot_lambda, aes(x = Year_t, y = lambda)) +
  geom_boxplot() +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  geom_line() +
  facet_wrap(~Population, labeller = as_labeller(PopFacet)) +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  theme_classic() +
  theme(strip.background = element_blank()) +
  # adds axes to each facet
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  # adjust facet labels
  theme(strip.text = element_text( size = 12, color = "black")) +
  theme(strip.text = element_text(hjust = 0)) +
  #font size
  theme(text = element_text(size = 14))  

# with  ggarrange
p <- ggplot(boot_lambda %>%
              filter(Population == "Bayfield"), aes(x = Year_t, y = lambda)) +
  geom_boxplot() +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  geom_line() +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  scale_y_continuous(limits=c(0.85,1.62)) +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())
p2 <- ggplot(boot_lambda %>%
              filter(Population == "Fanshawe"), aes(x = Year_t, y = lambda)) +
  geom_boxplot() +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  geom_line() +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  scale_y_continuous(limits=c(0.85,1.62)) +
  theme_classic() +
  theme(axis.title = element_blank(),
        axis.text = element_blank())
p3 <- ggplot(boot_lambda %>%
              filter(Population == "Ilderton"), aes(x = Year_t, y = lambda)) +
  geom_boxplot() +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  geom_line() +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  scale_y_continuous(limits=c(0.7,1.8)) +
  theme_classic()
p4 <- ggplot(boot_lambda %>%
              filter(Population == "RBG"), aes(x = Year_t, y = lambda)) +
  geom_boxplot() +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  geom_line() +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  scale_y_continuous(limits=c(0.7,1.8)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank())

Graphs <- ggarrange(p,  p2,  p3,  p4,
                      labels = c("(a)", "(b)", "(c)", "(d)"),
                    widths = c(1,1,1,1),
                      font.label = list(size = 11), 
                      hjust = c(-3.5,-0.7,-3.5, -0.7),
                      #hjust = c(-1.5,-0.2,-0.2, -1.5, -0.2, -0.2),
                      ncol = 2, nrow = 2, align = "h")
Graphs

# Mean and 95% CI
## Used for comparison with IPopM
data1 <- boot_lambda %>%
  group_by(Population, Year_t) %>%
  summarise(mean = mean(lambda),
            Q2.5 = quantile(lambda, probs = 0.025),
            Q97.5 = quantile(lambda, probs = 0.975))

p <- ggplot(data1, aes(x = Year_t, y = mean)) +
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), width = 0.1) +
    geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  geom_line() +
  facet_wrap(~Population) +
  geom_point() +
  theme_classic()
p

# Save graph
# ggsave(plot = Graphs, dpi = 600, width = 8, height = 6,
#       filename = "Population_Lambda.png",
#       path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = Graphs, dpi = 600, width = 8, height = 6,
#       filename = "Population_Lambda.pdf",
#       path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
```

# Deterministic Model

Gather parameters
```{r Parameters - Deterministic}
# First need year and population numbers
Year_t <- c(1,2,3)
Population <- c(1,2,3,4)

# Survival
S.Intercepts <- c(Bs.intercept, Fs.intercept, Is.intercept, Rs.intercept)
S.Slopes1 <- c(Bs.slope1, Fs.slope1, Is.slope1, Rs.slope1)

data1 <- as.data.frame(cbind(S.Intercepts, S.Slopes1, Population))

surv.d <- data.frame()
row_index <- 1
for(j in 1:length(Population)){
  for (i in 1:length(Year_t)) {
    surv.d[row_index, "s.int"] <-  data1[j, 1]# + data1[j, 3] * Year_t[i]
    surv.d[row_index, "Year_t"] <- Year_t[i]
    surv.d[row_index, "Population"] <- Population[j]
    # move to the next row
    row_index <- row_index + 1
    }
}

S.Slopes1 <- as.data.frame(cbind(S.Slopes1, Population))
surv.d <- left_join(surv.d, S.Slopes1)

# Growth
G.Intercepts <- c(Bg.intercept, Fg.intercept, Ig.intercept, Rg.intercept)
G.Slopes1 <- c(Bg.slope1, Fg.slope1, Ig.slope1, Rg.slope1)
G.Slopes2 <- c(Bg.slope2, Fg.slope2, Ig.slope2, Rg.slope2)
G.Slopes3 <- c(Bg.slope3, Fg.slope3, Ig.slope3, Rg.slope3)

data1 <- as.data.frame(cbind(G.Intercepts, G.Slopes1, G.Slopes2, G.Slopes3, Population))

grow.d <- data.frame()
row_index <- 1
for(j in 1:length(Population)){
  for (i in 1:length(Year_t)) {
    grow.d[row_index, "g.int"] <-  data1[j, 1] + data1[j, 2] * Year_t[i]
    grow.d[row_index, "Year_t"] <- Year_t[i]
    grow.d[row_index, "Population"] <- Population[j]
    # move to the next row
    row_index <- row_index + 1
    }
}

G.Slopes2 <- as.data.frame(cbind(G.Slopes2, Population))
G.Slopes3 <- as.data.frame(cbind(G.Slopes3, Population))
grow.d <- left_join(grow.d, G.Slopes2)
grow.d <- left_join(grow.d, G.Slopes3)

# Growth Variance
growth_var <- data.frame(Grow_sd = double(), Population = double())
growth_var[1,1] <- Bgrowth_var
growth_var[2,1] <- Fgrowth_var
growth_var[3,1] <- Igrowth_var
growth_var[4,1] <- Rgrowth_var
growth_var[1,2] <- 1 # Bayfield
growth_var[2,2] <- 2 # Fanshawe
growth_var[3,2] <- 3 # Ilderton
growth_var[4,2] <- 4 # RBG

# Flowers
F.Intercepts <- c(Bf.intercept, Ff.intercept, If.intercept, Rf.intercept)
F.Slopes1 <- c(Bf.slope1, Ff.slope1, If.slope1, Rf.slope1)
F.Slopes2 <- c(Bf.slope2, Ff.slope2, If.slope2, Rf.slope2)
F.Slopes3 <- c(Bf.slope3, Ff.slope3, If.slope3, Rf.slope3)

data1 <- as.data.frame(cbind(F.Intercepts, F.Slopes1, F.Slopes2, F.Slopes3, Population))

flower.d <- data.frame()
row_index <- 1
for(j in 1:length(Population)){
  for (i in 1:length(Year_t)) {
    flower.d[row_index, "fl.int"] <-  data1[j, 1] + data1[j, 2] * Year_t[i]
    flower.d[row_index, "Year_t"] <- Year_t[i]
    flower.d[row_index, "Population"] <- Population[j]
    # move to the next row
    row_index <- row_index + 1
    }
}

F.Slopes2 <- as.data.frame(cbind(F.Slopes2, Population))
F.Slopes3 <- as.data.frame(cbind(F.Slopes3, Population))
flower.d <- left_join(flower.d, F.Slopes2)
flower.d <- left_join(flower.d, F.Slopes3)

#Seeds
Se.Intercepts <- c(Bse.intercept, Fse.intercept, Ise.intercept, Rse.intercept)
Se.Slopes1 <- c(Bse.slope1, Fse.slope1, Ise.slope1, Rse.slope1)
Se.Slopes2 <- c(Bse.slope2, Fse.slope2, Ise.slope2, Rse.slope2)
Se.Slopes3 <- c(Bse.slope3, Fse.slope3, Ise.slope3, Rse.slope3)

data1 <- as.data.frame(cbind(Se.Intercepts, Se.Slopes1, Se.Slopes2, Se.Slopes3, Population))

seed.d <- data.frame()
row_index <- 1
for(j in 1:length(Population)){
  for (i in 1:length(Year_t)) {
    seed.d[row_index, "se.int"] <-  data1[j, 1] + data1[j, 2] * Year_t[i]
    seed.d[row_index, "Year_t"] <- Year_t[i]
    seed.d[row_index, "Population"] <- Population[j]
    # move to the next row
    row_index <- row_index + 1
    }
}

Se.Slopes2 <- as.data.frame(cbind(Se.Slopes2, Population))
Se.Slopes3 <- as.data.frame(cbind(Se.Slopes3, Population))
seed.d <- left_join(seed.d, Se.Slopes2)
seed.d <- left_join(seed.d, Se.Slopes3)


### Combine them all into one dataframe
Params <- left_join(surv.d, grow.d)
Params <- left_join(Params, flower.d, by = c("Population", "Year_t"))
Params <- left_join(Params, seed.d, by = c("Population", "Year_t"))

# Rename slopes to match boot_params
colnames(Params)[colnames(Params) == 'S.Slopes1'] <- 's.slope'
colnames(Params)[colnames(Params) == 'G.Slopes2'] <- 'g.slope'
colnames(Params)[colnames(Params) == 'G.Slopes3'] <- 'g.slope2'
colnames(Params)[colnames(Params) == 'F.Slopes2'] <- 'f.slope'
colnames(Params)[colnames(Params) == 'F.Slopes3'] <- 'f.slope2'

colnames(Params)[colnames(Params) == 'Se.Slopes2'] <- 'se.slope'
colnames(Params)[colnames(Params) == 'Se.Slopes3'] <- 'se.slope2'

### Add other things that aren't bootstrapped--these are considered to be the
# same across all populations and years
Params$Prob_Germ_1 <- Prob_Germ_1
Params$Prob_Germ_2 <- Prob_Germ_2
Params$Prob_Surv_Seed <- Prob_Surv_Seed
Params$Estab <- est.intercepts
Params$Recruit_size <- Recr_mean$Recr_mean[match(Params$Population, Recr_mean$Population)]
Params$Recruit_sd <- Recr_sd$Recr_sd[match(Params$Population, Recr_sd$Population)]

# Add the growth variation--this is considered to be the same across years
#  but can be different between the populations
Params$grow.sd <- growth_var$Grow_sd[match(Params$Population, growth_var$Population)]

# Make into a list, one element for each row
params_list <-  filter(Params) %>%
  split(., seq(nrow(.)))
```

## Functions
```{r Functions - Deterministic}
########## Eviction ########################################
# Avoid unintentional eviction
#Set minimum and maximum sizes
min.size = 0.9 * min(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
max.size = 1.1 * max(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)

x_range <- function(Height_cm_t, Params) {
  pmin(pmax(Height_cm_t, min.size), max.size)
}

######### Survival #########################################
#Survival of established plants
S_z <- function(Height_cm_t, Year_t, Params) {
  xb <- x_range(Height_cm_t, Params)
  linear_p <- Params$s.int + Params$s.slope * xb
  p <- 1/(1 + exp(-linear_p))
  return(p)
}

#Test this
mean(S_z(12, 1, Params))
mean(S_z(70, 2, Params))
#Seems ok

######### Growth ###########################################

G_z1z <- function(Height_cm_t, Height_cm_t1, Year_t, Params) {
  xb <- x_range(Height_cm_t, Params)
  return(dnorm(Height_cm_t1,
  mean <- Params$g.int + Params$g.slope * xb + Params$g.slope2 * Year_t * xb,
  sd = Params$grow.sd))
}

#Test this
mean(G_z1z(12, 15, 1, Params))
mean(G_z1z(62, 70, 2, Params))
# I think this is right

######### Combine Survival and Growth ######################
P_z <- function(Height_cm_t, Height_cm_t1, Year_t, Params){
  xb <- x_range(Height_cm_t, Params)
  return(S_z(xb, Year_t, Params) * G_z1z(xb, Height_cm_t1, Year_t, Params))
}

#Test this
mean(P_z(12, 15, 1, Params))
mean(P_z(54, 72, 2, Params))

######### Probability of Flowering #########################

#based on year t
Pf_z <- function(Height_cm_t, Year_t, Params) {
  xb <- x_range(Height_cm_t, Params)
  u <- exp(Params$fl.int + Params$f.slope * xb + Params$f.slope2 * Year_t * xb)
  p <- 1/(1 + exp(-u))
  return(p)
}

#Test this
mean(Pf_z(12, 1, Params))
mean(Pf_z(72, 2, Params))
# Yup

######### Seeds Per Plant ##################################

Sf_z <- function(Height_cm_t, Year_t, Params){
  nseed <- Params$se.int + Params$se.slope * Height_cm_t + Params$se.slope2 * Year_t * Height_cm_t
  return(exp(nseed)) # The seeds values are logged, so I need to reverse them to real values
}

#Test this
mean(Sf_z(12, 1, Params))
mean(Sf_z(72, 2, Params))
# Yup!!

######### Flowering Probability and Seed Number ############

F_z <- function(Height_cm_t, Year_t, Params) {
  xb <- x_range(Height_cm_t, Params)
  x <- Pf_z(xb, Year_t, Params) * Sf_z(xb, Year_t, Params)
  return(x)
}

# Test this
mean(F_z(12, 1, Params))
mean(F_z(72, 2, Params))

########## Seedling Sizes ##################################
Se_z <- function(Height_cm_t1, Params){
  dnorm(x = Height_cm_t1,
        mean = Params$Recruit_size,
        sd = Params$Recruit_sd)
}

#Test this
mean(Se_z(5, Params))
mean(Se_z(10, Params))

########## Seedling Survival ###############################
Sur_z1 <- function(Params){
  surv_adults <- Params$Estab
  return(surv_adults)
}

#Test this
mean(Sur_z1(Params))
#yep

########## Seed bank #######################################
S1_z <- function(Height_cm_t, Year_t,Params){
  f <- F_z(Height_cm_t, Year_t, Params) * (1 - Params$Prob_Germ_1)
  return(f);
}

#Test this
mean(S1_z(12, 1, Params))
mean(S1_z(72, 2, Params))
#looks pretty good
```

## Kernel
```{r Kernel - Deterministic, warning=FALSE}
#Set minimum and maximum sizes
min.size <- 0.9 * min(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
max.size <- 1.1 * max(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)

#building off of code from Tom Miller:
mat.size = 100 #dimensions of continuous matrix

bigmatrix <- function(params_list, mat.size){
  n <- mat.size #matrix size
  L <- min.size #lower integration limit
  U <- max.size #upper integration limit
  
  h <- (U - L)/n #Bin size
  b <- L + c(0:n) * h #Lower boundaries 
  y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
  # y is the mesh points of the matrix
  
  # Fertility matrix
  Fmat<-matrix(0,(n+4),(n+4))
  
  # Fecundity
  
  #Seedlings this year
  Fmat[4,5:(n+4)] <- Prob_Germ_1 * F_z(y, Year_t, params_list) + #seedlings this years seed
    Prob_Germ_2 * Prob_Surv_Seed * S1_z(y, Year_t, params_list) + #last year seed
    Prob_Germ_2 * (1 - Prob_Germ_1) * Prob_Surv_Seed * Prob_Germ_2 * Prob_Surv_Seed *  S1_z(y, Year_t, params_list) #2 year ago seed
  
  #Seedbank with t-1 and t-2 seeds
  Fmat[1,5:(n+4)] <- (1 - Prob_Germ_1) * (1 - Prob_Germ_2) * Prob_Surv_Seed * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #3-year old seeds
  Fmat[2,5:(n+4)] <- (1 - Prob_Germ_1) * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #2-year old seeds
  Fmat[3,5:(n+4)] <- (1 - Prob_Germ_1 - Prob_Germ_2) * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #1-year old seeds
  
  # Growth/survival transition matrix
  Tmat<-matrix(0,(n+4),(n+4))
  
  # probability a seed stays in the seedbank
  Tmat[2,1] <- (1 - Prob_Germ_1) * Prob_Surv_Seed
  Tmat[3,1] <- (1 - Prob_Germ_1) * Prob_Surv_Seed * (1 - Prob_Germ_2) * Prob_Surv_Seed
  Tmat[3,2] <- (1 - Prob_Germ_1) * Prob_Surv_Seed
  
  # probability seed germinates after 2 years
  Tmat[4,1] <- Prob_Germ_2 * Prob_Surv_Seed * (1 - Prob_Germ_2) * Prob_Surv_Seed
  Tmat[4,2] <- Prob_Germ_2 * Prob_Surv_Seed
  
  # probability seed germinates fresh
  Tmat[4,3] <- Prob_Germ_1
  
  # probability a seedling establishes as adult
  Tmat[5:(n+4),4] <- Sur_z1(params_list) * Se_z(y, params_list)
  
  # survival and growth
  Tmat[5:(n+4),5:(n+4)] <- (outer(y,y,P_z, Year_t, params_list)) * h #removed transpose of outer()
  
  # Full kernel -- adds two kernels built
  IPMmat <- Fmat + Tmat
  
  out_matrix <- list(IPMmat = IPMmat, Fmat = Fmat, Tmat = Tmat, meshpts = y)

  return(out_matrix)
}

# Run for all rows of Params
lamb_lists <- lapply(X = params_list, FUN = bigmatrix, mat.size = mat.size)
```


Save matrices
```{r eval=FALSE, include=FALSE}
#save model output
# saveRDS(lamb_lists, file = "~/Research/Data/IPM_Formatted_STYL/Output/IPM_Matrix.rds")

# Load IPM output
# lamb_lists <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPM_Matrix.rds")
```

Check lambda
```{r}
### Lambda
GrowthData <- data.frame(matrix(ncol = 3, nrow = 12))
colnames(GrowthData) <- c("Population", "Year_t", "lambda")
GrowthData$Year_t <- as.factor(c(rep(c(2022, 2023, 2024), 4)))
GrowthData$Population <- c("Bayfield", "Bayfield", "Bayfield",
                           "Fanshawe", "Fanshawe", "Fanshawe",
                           "Ilderton", "Ilderton", "Ilderton",
                           "RBG", "RBG", "RBG")

GrowthData[1,3] <- lambda(lamb_lists$`1`$IPMmat)
GrowthData[2,3] <- lambda(lamb_lists$`2`$IPMmat)
GrowthData[3,3] <- lambda(lamb_lists$`3`$IPMmat)
GrowthData[4,3] <- lambda(lamb_lists$`4`$IPMmat)
GrowthData[5,3] <- lambda(lamb_lists$`5`$IPMmat)
GrowthData[6,3] <- lambda(lamb_lists$`6`$IPMmat)
GrowthData[7,3] <- lambda(lamb_lists$`7`$IPMmat)
GrowthData[8,3] <- lambda(lamb_lists$`8`$IPMmat)
GrowthData[9,3] <- lambda(lamb_lists$`9`$IPMmat)
GrowthData[10,3] <- lambda(lamb_lists$`10`$IPMmat)
GrowthData[11,3] <- lambda(lamb_lists$`11`$IPMmat)
GrowthData[12,3] <- lambda(lamb_lists$`12`$IPMmat)

GrowthData
```

# Basic analysis

Code from:
https://github.com/ipmbook/first-edition/blob/master/Rcode/c4/Ungulate%20Pert%20Lambda%20Calc.R

## Kernel Elasticity

```{r}
#Set minimum and maximum sizes
min.size <- 0.9 * min(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
max.size <- 1.1 * max(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)

mat.size <- 100 #dimensions of continuous matrix

lambda <-  data.frame(matrix(ncol = 12, nrow = 1))
#provide column names
colnames(lambda) <- c('Bayfield1', 'Fanshawe1', 'Ilderton1', 'RBG1','Bayfield2', 'Fanshawe2', 'Ilderton2', 'RBG2',
                      'Bayfield3', 'Fanshawe3', 'Ilderton3', 'RBG3')
w.z <- data.frame(matrix(ncol = 12, nrow = 104))
#provide column names
colnames(w.z) <- c('Bayfield1', 'Fanshawe1', 'Ilderton1', 'RBG1','Bayfield2', 'Fanshawe2', 'Ilderton2', 'RBG2',
                      'Bayfield3', 'Fanshawe3', 'Ilderton3', 'RBG3')
v.z1 <- data.frame(matrix(ncol = 12, nrow = 104))
#provide column names
colnames(v.z1) <- c('Bayfield1', 'Fanshawe1', 'Ilderton1', 'RBG1','Bayfield2', 'Fanshawe2', 'Ilderton2', 'RBG2',
                      'Bayfield3', 'Fanshawe3', 'Ilderton3', 'RBG3')

stable.dist <- data.frame(matrix(ncol = 12, nrow = 104))
#provide column names
colnames(stable.dist) <- c('Bayfield1', 'Fanshawe1', 'Ilderton1', 'RBG1','Bayfield2', 'Fanshawe2', 'Ilderton2', 'RBG2',
                      'Bayfield3', 'Fanshawe3', 'Ilderton3', 'RBG3')
repro.val <- data.frame(matrix(ncol = 12, nrow = 104))
#provide column names
colnames(repro.val) <- c('Bayfield1', 'Fanshawe1', 'Ilderton1', 'RBG1','Bayfield2', 'Fanshawe2', 'Ilderton2', 'RBG2',
                      'Bayfield3', 'Fanshawe3', 'Ilderton3', 'RBG3')
#make an empty list
K <- vector(mode='list', length = 12)
P <- vector(mode='list', length = 12)
F <- vector(mode='list', length = 12)

K.sens <- vector(mode='list', length = 12)
K.elas <- vector(mode='list', length = 12)

P.elas <- vector(mode='list', length = 12)
F.elas <- vector(mode='list', length = 12)

#keep track of if things add to 1 as they should
K.test <- vector(mode='list', length = 12)

##############################
  n <- mat.size  #matrix size
  L <- min.size #lower integration limit
  U <- max.size #upper integration limit
 
  h <- (U - L)/n #Bin size
  b <- L + c(0:n) * h #Lower boundaries 
  y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
  # y is the mesh points of the matrix

for(i in 1:12){
  IPM.sys <- bigmatrix(params_list[[i]],100)
  
  ## extract the mesh points and the kernel
  meshpts <- IPM.sys$meshpts
  h <- diff(meshpts[1:2])
  K[[i]] <- IPM.sys$IPMmat
  P[[i]] <- IPM.sys$Tmat
  F[[i]] <- IPM.sys$Fmat
  
  ## compute the eigen vectors / values 
  IPM.eig.sys <- eigen(K[[i]])
  ## lambda
  lambda[,i] <- Re(IPM.eig.sys$values[1])
  ## unnormalised stable size distribution ('w')
  w.z[,i] <- Re(IPM.eig.sys$vectors[,1])
  
  ## unnormalised reproductive value distribution ('v')
  v.z1[,i] <- Re(eigen(t(K[[i]]))$vectors[,1])
  ################
  
  ## kernel sensitivity
  K.sens[[i]] <- outer(v.z1[,i], w.z[,i], "*")/sum(v.z1[,i] * w.z[,i] * h)
  ## kernel elasticity
  K.elas[[i]] <- K.sens[[i]] * (K[[i]]/h) / lambda[,i]
  ## does the kernel elasticity integrate to 1?
  K.test[i] <- sum(K.elas[[i]]) * h^2

  ## calculate the elasticities of the survival and reproduction components
  P[[i]] <- IPM.sys$Tmat  / h
  F[[i]]  <- IPM.sys$Fmat  / h
  P.elas[[i]] <- P[[i]]  * K.sens[[i]] / lambda[,i]
  F.elas[[i]] <- F[[i]]  * K.sens[[i]] / lambda[,i]
  
  ## get the relative contributions of survival and reproduction to lambda
  sum(P.elas[[i]]) * h^2
  sum(F.elas[[i]]) * h^2
}

# Check that elasticities sum to one before running anything else
K.test
# They are!

######################################
#plot sensitivity
  n <- mat.size# + 4  #matrix size
  L <- min.size #lower integration limit
  U <- max.size #upper integration limit
  h <- (U - L)/n #Bin size
  b <- L + c(0:n) * h #Lower boundaries 
  y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
  # y is the mesh points of the matrix
  
# plot one of these
#png(filename = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs/Elasticity.png",
#    width = 6, height = 6, units = "in", res = 600)
  
data1 <- K.elas[[1]]

# make separate image for seedlings
data2 <- data1
data2[5:104,5:104] <- NA
data2[1:3, 1:3] <- NA
y2 <- c(-0.1, -0.01, -0.001, 0, y)

# and seedbank
data3 <- data1
data3[4:104,4:104] <- NA
y3 <- c(-3, -2, -1, -0.99, y)


image(y,y,t(data1[5:104,5:104]), xlab="Size Year t",ylab="Size Year t+1", main="Elasticity",
           col = terrain.colors(30, rev = T), xlim = c(-3, 90), ylim = c(-3,90))

image(y2,y2,t(data2),
           col = terrain.colors(30, rev = T), add = T)

image(y3,y3,t(data3),
           col = terrain.colors(30, rev = T), add = T)
#dev.off()

#### Another version with a scale! #####

# plot one of these
#png(filename = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs/Elasticity_scale.png",
#    width = 7, height = 6, units = "in", res = 600)

# zlim makes them the same scale
image.plot(y,y,t(data1[5:104,5:104]), xlab="Size Year t",ylab="Size Year t+1",
           col = terrain.colors(30, rev = T), xlim = c(-3, 90), ylim = c(-3,90), zlim = c(0,0.007))

image.plot(y2,y2,t(data2)^1, zlim = c(0,.007),
           col = terrain.colors(30, rev = T), add = T)

image.plot(y3,y3,t(data3)^1, zlim = c(0,0.007),
           col = terrain.colors(30, rev = T), add = T)
#dev.off()
```

Make figure with all kernels for supplemental
```{r}
#plot sensitivity
  n <- mat.size + 4  #matrix size
  L <- min.size #lower integration limit
  U <- max.size #upper integration limit
  h <- (U - L)/n #Bin size
  b <- L + c(0:n) * h #Lower boundaries 
  y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
  # y is the mesh points of the matrix
  
#png(filename = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs/Elasticity_allYrs.png",
#    width = 8, height = 6, units = "in", res = 600)

par(mfrow = c(4,3)) # so each row is a population

# also need to modify the plot margins for each plot

par(mar=c(1,4,1,1))
image(y,y,t(K.elas[[1]]),xlab="",ylab="Size (t+1)", col = terrain.colors(30, rev = T), xaxt='n')
#text(45,84, paste("2022-2023"))
text(11,85, paste("Bayfield"))
par(mar=c(1,1,1,1))
image(y,y,t(K.elas[[2]]),xlab="",ylab="", col = terrain.colors(30, rev = T), xaxt='n', yaxt="n")
#text(45,84, paste("2023-2024"))
par(mar=c(1,1,1,1))
image(y,y,t(K.elas[[3]]),xlab="",ylab="", col = terrain.colors(30, rev = T), xaxt='n', yaxt="n")
#text(45,84, paste("2024-2025"))

par(mar=c(1,4,1,1))
image(y,y,t(K.elas[[4]]),xlab="",ylab="Size (t+1)", col = terrain.colors(30, rev = T), xaxt='n')
text(13,85, paste("Fanshawe"))
par(mar=c(1,1,1,1))
image(y,y,t(K.elas[[5]]),xlab="",ylab="", col = terrain.colors(30, rev = T), xaxt='n', yaxt="n")
par(mar=c(1,1,1,1))
image(y,y,t(K.elas[[6]]),xlab="",ylab="", col = terrain.colors(30, rev = T), xaxt='n', yaxt="n")

par(mar=c(1,4,1,1))
image(y,y,t(K.elas[[7]]),xlab="",ylab="Size (t+1)", col = terrain.colors(30, rev = T), xaxt='n')
text(11,85, paste("Ilderton"))
par(mar=c(1,1,1,1))
image(y,y,t(K.elas[[8]]),xlab="",ylab="", col = terrain.colors(30, rev = T), xaxt='n', yaxt="n")
par(mar=c(1,1,1,1))
image(y,y,t(K.elas[[9]]),xlab="",ylab="", col = terrain.colors(30, rev = T), xaxt='n', yaxt="n")

par(mar=c(4,4,1,1))
image(y,y,t(K.elas[[10]]),xlab="Size (t)",ylab="Size (t+1)", col = terrain.colors(30, rev = T))
text(8,85, paste("RBG"))
par(mar=c(4,1,1,1))
image(y,y,t(K.elas[[11]]),xlab="Size (t)",ylab="", col = terrain.colors(30, rev = T), yaxt="n")
par(mar=c(4,1,1,1))
image(y,y,t(K.elas[[12]]),xlab="Size (t)",ylab="", col = terrain.colors(30, rev = T), yaxt="n")

#dev.off()
```

## Stable Size Distribution

```{r}
stable.dist <- data.frame(matrix(ncol = 12, nrow = 104))

#plot sensitivity
n <- mat.size   #matrix size
L <- min.size #lower integration limit
U <- max.size #upper integration limit
h <- (U - L)/n #Bin size
b <- L + c(0:n) * h #Lower boundaries 
y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
# y is the mesh points of the matrix

# Normalize the stable distribution
for(i in 1:12){
  stable.dist[,i] <- w.z[1:104,i]/sum(w.z[1:104,i])
}

#maybe don't use smallest sizes
y <- c(0.5, 1, 1.5, 2, y)

# GGPlot version
names(stable.dist) <- c("B1", "B2","B3", "F1","F2","F3","I1","I2", "I3","R1","R2", "R3")
data4 <- stable.dist %>%
  mutate(y = y) %>%
  pivot_longer(cols = c("B1", "B2","B3", "F1","F2","F3","I1","I2", "I3","R1","R2", "R3"), names_to = "Population", values_to = "StableSize") %>%
  mutate(Year = case_when(grepl(1, Population) ~ 2022,
                          grepl(2, Population) ~ 2023,
                          grepl(3, Population) ~ 2024)) %>%
  mutate(Population = case_when(grepl("B", Population) ~ "Bayfield",
                                grepl("F", Population) ~ "Fanshawe",
                                grepl("I", Population) ~ "Ilderton",
                                grepl("R", Population) ~ "RBG"))
data4$Year <- as.factor(data4$Year)

p <- ggplot(data=data4 %>%
              filter(Population == "Bayfield") %>%
              filter(y > 1.5), aes(x=y, y=StableSize, group=Year)) +
  geom_line(aes(linetype = Year), col = "#D55E00", size = 1.2) +
  labs(y = "Relative Size Distribution", x = "Height (cm), Year t") +
  theme_classic() +
  # Increase font size of population names
  theme(strip.text = element_text(size = 16)) +
    facet_wrap(~Year) +
  #add axes on all graphs, instead of just bottom and left ones
  theme(axis.line=element_line()) +
  scale_x_continuous(limits=c(0,90)) +
  scale_y_continuous(limits=c(0,0.3)) +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank()) +
  #move the legend
  theme(legend.position = "none") +
  theme(strip.background = element_blank())

p2 <- ggplot(data=data4 %>%
              filter(Population == "Fanshawe") %>%
              filter(y > 1.5), aes(x=y, y=StableSize, group=Year)) +
  geom_line(aes(linetype = Year), col = "#0072B2",  size = 1.2) +
  labs(y = "Relative Size Distribution", x = "Height (cm), Year t") +
  theme_classic() +
  # Increase font size of population names
  theme(strip.text = element_text(size = 16)) +
    facet_wrap(~Year) +
  #add axes on all graphs, instead of just bottom and left ones
  theme(axis.line=element_line()) +
  scale_x_continuous(limits=c(0,90)) +
  scale_y_continuous(limits=c(0,0.3)) +
  theme(axis.title = element_blank(),
        axis.text=element_blank()) +
  #move the legend
  theme(legend.position = "none") +
  theme(strip.background = element_blank())

p3 <- ggplot(data=data4 %>%
              filter(Population == "Ilderton") %>%
              filter(y > 1.5), aes(x=y, y=StableSize, group=Year)) +
  geom_line(aes(linetype = Year), col = "#F0E442", size = 1.2) +
  labs(y = "Relative Size Distribution", x = "Height (cm), Year t") +
  theme_classic() +
  # Increase font size of population names
  theme(strip.text = element_text(size = 16)) +
    facet_wrap(~Year) +
  #add axes on all graphs, instead of just bottom and left ones
  theme(axis.line=element_line()) +
  scale_x_continuous(limits=c(0,90)) +
  scale_y_continuous(limits=c(0,0.3)) +
  #move the legend
  theme(legend.position = "none") +
  theme(strip.background = element_blank())

p4 <- ggplot(data=data4 %>%
              filter(Population == "RBG") %>%
              filter(y > 1.5), aes(x=y, y=StableSize, group=Year)) +
  geom_line(aes(linetype = Year), col = "#009E73", size = 1.2) +
  labs(y = "Relative Size Distribution", x = "Height (cm), Year t") +
  theme_classic() +
  # Increase font size of population names
  theme(strip.text = element_text(size = 16)) +
    facet_wrap(~Year) +
  #add axes on all graphs, instead of just bottom and left ones
  theme(axis.line=element_line()) +
  scale_x_continuous(limits=c(0,90)) +
  scale_y_continuous(limits=c(0,0.3)) +
  theme(axis.title.y = element_blank(),
        axis.text.y=element_blank()) +
  # move the legend
  theme(legend.position = "none") +# c(0.85, 0.40)) +
  theme(strip.background = element_blank())

Graphs <- ggarrange(p,  p2,  p3,  p4,
                      labels = c("(a) Bayfield", "(b) Fanshawe", "(c) Ilderton", "(d) RBG"),
                    widths = c(1,1,1,1),
                      font.label = list(size = 11), 
                      hjust = c(-0.5,-0.2, -0.7, -0.3),#c(-1.5,-0.7,-1.5, -0.7),
                    vjust = c(0.2, 0.2, 0.2, 0.2),
                      #hjust = c(-1.5,-0.2,-0.2, -1.5, -0.2, -0.2),
                      ncol = 2, nrow = 2, align = "h") +
  # add extra margin above all plots
  theme(plot.margin = margin(1, 0, 0,0, "cm"))

Graphs

# save plot
 ggsave(plot = Graphs, dpi = 600, width = 8, height = 6,
        filename = "StableSizeDist.png",
        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = Graphs, dpi = 600, width = 8, height = 6,
#        filename = "StableSizeDist.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
```

# Stochastic Growth Rates

Code from:
https://github.com/ipmbook/first-edition/blob/master/Rcode/c7/Carlina/Carlina%20Fixed%20Effects%20K%20pert.R

```{r Stochastic lambda}
#Set minimum and maximum sizes
min.size <- 0.9 * min(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
max.size <- 1.1 * max(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)

mat.size <- 100 #dimensions of continuous matrix

nBigMatrix <- 100
n.est <- 20000
n.runin <- 500
min.size = 0.9 * min(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
max.size = 1.1 * max(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
n.years <- 3


n <- mat.size #matrix size
L <- min.size #lower integration limit
U <- max.size #upper integration limit
h <- (U - L)/n #Bin size
b <- L + c(0:n) * h #Lower boundaries 
y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints

stoc_pert_analysis<-function(Params,n.est,n.runin){
	
	year.i <- sample(1:n.years,n.est+1,replace=TRUE)

	K.year.i <- array(NA,c((nBigMatrix+4),(nBigMatrix+4),n.years))
	
	for(i in 1:n.years){
		year.K<-bigmatrix(Params[i,],nBigMatrix)
		K.year.i[,,i] <- year.K$IPMmat
	}

#Calculate mean kernel, v and w
	mean.kernel <- apply(K.year.i[,,i],1:2,mean)
	
	w <- Re(eigen(mean.kernel)$vectors[,1]); 
	v <- Re(eigen(t(mean.kernel))$vectors[,1]);

	# scale eigenvectors <v,w>=1 
	w <- abs(w)/sum(h*abs(w))
	v <- abs(v)
	v <- v/(h*sum(v*w))
    cat(h*sum(v*w)," should = 1","\n")

#Esimate Lambda s
	nt<-rep(1/(nBigMatrix+4),(nBigMatrix+4))
	rt.V <- rt.N <- rep(NA,n.est)
	
#Iterate model
	for (year.t in 1:n.est){
		if(year.t%%10000==0) cat("iterate: ", year.t,"\n");

		#iterate model with year-specific kernel
		nt1<-K.year.i[,,year.i[year.t]] %*% nt
	
		sum.nt1<-sum(nt1)
		
		#Calculate log growth rates  
		rt.V[year.t] <- log(sum(nt1*v)/sum(nt*v))
		rt.N[year.t] <- log(sum(nt1)/sum(nt))
		nt <- nt1 / sum.nt1  
	}

Ls <- exp(mean(rt.V))
	
### Get wt and Rt time series ###
	wt<-matrix(1/nBigMatrix, nrow=n.est+1, ncol=(nBigMatrix+4));
	for (i in 1:n.est) {
		
		K             <- K.year.i[,,year.i[i]]
		wt[i+1,]  <-K %*% wt[i,]
		wt[i+1,]  <-wt[i+1,]/sum(wt[i+1,]);
		if(i%%10000==0) cat("wt ",i,"\n")

	}

### Get vt time series ###
	vt<-matrix(1/nBigMatrix, nrow=n.est+1, ncol=(nBigMatrix+4));
	for (i in (n.est+1):2) {
		K           <- K.year.i[,,year.i[i]]
		vt[i-1,]  <- vt[i,] %*% K
		vt[i-1,]  <- vt[i-1,]/sum(vt[i-1,]);
		if(i%%10000==0) cat("vt  ",i,"\n")

	}

sens.s <- matrix(0,nrow=(nBigMatrix+4),ncol=(nBigMatrix+4));
elas.s <- sens.s

for (i in n.runin:(n.est-n.runin)) {
	
		K                 <-   K.year.i[,,year.i[i]]
		vt1.wt        <-   vt[i+1,]%*%t(wt[i,])
		vt1.K.wt   <- sum(vt[i+1,] * (K %*% wt[i,]))

		#calculation of the standard sensitivities and elasticities
	
		sens.s<-sens.s+vt1.wt/vt1.K.wt;

		elas.s<-elas.s+K*(vt1.wt/vt1.K.wt);
		
}

 elas.s <- elas.s/(n.est-2*n.runin+1)
 sens.s <- Ls*sens.s/(n.est-2*n.runin+1)
 
 return(list(meshpts=year.K$meshpts,h=h,sens.s=sens.s,elas.s=elas.s,mean.kernel=mean.kernel,Ls=Ls))

}

#  I need to run this function on each population separately
# Each population has 3 rows in the Params dataframe
# I'll just hard-code it
#Bayfield
Params2 <- as.data.frame(Params[1:3,])
Bpert.K <- stoc_pert_analysis(Params2,n.est,n.runin)
#Fanshawe
Params2 <- as.data.frame(Params[4:6,])
Fpert.K <- stoc_pert_analysis(Params2,n.est,n.runin)
#Ilderton
Params2 <- as.data.frame(Params[7:9,])
Ipert.K <- stoc_pert_analysis(Params2,n.est,n.runin)
#RBG
Params2 <- as.data.frame(Params[10:12,])
Rpert.K <- stoc_pert_analysis(Params2,n.est,n.runin)

####
Stoch_lambda <-  data.frame(matrix(nrow = 4, ncol = 2))
colnames(Stoch_lambda) <- c("Population", "Lambda")
Stoch_lambda[,1] <- c("Bayfield", "Fansahwe", "Ilderton", "RBG")
Stoch_lambda[1,2] <- Bpert.K$Ls
Stoch_lambda[2,2] <- Fpert.K$Ls
Stoch_lambda[3,2] <- Ipert.K$Ls
Stoch_lambda[4,2] <- Rpert.K$Ls

Stoch_lambda

# Save to a dataframe
# write.csv(Stoch_lambda, "~/Research/Data/IPM_Formatted_STYL/Output/Stoch_Lambda.csv", row.names = FALSE)

# Just to check that I did this right
# https://rdrr.io/cran/popbio/man/stoch.growth.rate.html
Blamb_lists <- list(lamb_lists$`1`$IPMmat, lamb_lists$`2`$IPMmat, lamb_lists$`3`$IPMmat)
testdata <- stoch.growth.rate(Blamb_lists, verbose = T)
exp(testdata$sim)
# Yup, this is nearly the same for Bayfield!
```

# Parameter-Level Elasticity

all code from https://github.com/jloesberg/Primula/blob/main/scripts/primula_IPM.rmd

```{r}
#goal: examine sensitivity of parameters
perturbation <- 0.01 #perturbation

sens.params <- Params %>%
  # Add seed number
  mutate(Seeds = case_when(Population == 1 ~ Bayfield_seeds,
                           Population == 2 ~ Fanshawe_seeds,
                           Population == 3 ~ Ilderton_seeds,
                           Population == 4 ~ RBG_seeds,
                           TRUE ~ NA_real_))

sens.params$version <- "original"

sens.params.g.int <- sens.params %>% 
  mutate(g.int = g.int + perturbation,
         version = "g.int")
sens.params.g.slope <- sens.params %>% 
  mutate(g.slope = g.slope + perturbation,
         g.slope2 = g.slope2 + perturbation,##
         version = "g.slope")
sens.params.surv.int <- sens.params %>% 
  mutate(s.int = s.int + perturbation,
         version = "s.int")
sens.params.surv.slope <- sens.params %>% 
  mutate(s.slope = s.slope + perturbation,
         version = "s.slope")
sens.params.pflow.int <- sens.params %>% 
  mutate(fl.int = fl.int + perturbation,
         version = "fl.int")
sens.params.pflow.slope <- sens.params %>% 
  mutate(f.slope = f.slope + perturbation,
         f.slope2 = f.slope + perturbation,##
         version = "f.slope")
sens.params.numse.int <- sens.params %>% 
  mutate(se.int = se.int + perturbation,
         version = "se.int")
sens.params.numse.slope <- sens.params %>% 
  mutate(se.slope = se.slope + perturbation,
         se.slope2 = se.slope2 + perturbation,##
         version = "se.slope")
sens.params.rec.size.mean <- sens.params %>% 
  mutate(Recruit_size = Recruit_size + perturbation,
         version = "Recruit_size")
sens.paramsProb_Germ_1 <- sens.params %>% 
  mutate(Prob_Germ_1 = Prob_Germ_1 + perturbation,
         version = "Prob_Germ_1")
sens.paramsProb_Germ_2 <- sens.params %>% 
  mutate(Prob_Germ_2 = Prob_Germ_2 + perturbation,
         version = "Prob_Germ_2")
sens.paramsSurvSeed <- sens.params %>% 
  mutate(Prob_Surv_Seed = Prob_Surv_Seed + perturbation,
         version = "Prob_Surv_Seed")

sens.paramsSurvEstab <- sens.params %>% 
  mutate(Estab = Estab + perturbation,
         version = "Estab")

# make list of models
sens <- mget(ls(pattern = "^sens.params"))

sens <- do.call("rbind", sens)

#split list by row
ltre_list <-  filter(sens) %>% split(., seq(nrow(.))) 

# Kernel

#Set minimum and maximum sizes
min.size <- 0.9 * min(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)
max.size <- 1.1 * max(c(SurvData$Height_cm_t, SurvData$Height_cm_t1), na.rm = T)

mat.size = 100 #dimensions of continuous matrix

bigmatrix<-function(ltre_list, mat.size){
  n <- mat.size #matrix size
  L <- min.size #lower integration limit
  U <- max.size #upper integration limit
  
  h <- (U - L)/n #Bin size
  b <- L + c(0:n) * h #Lower boundaries 
  y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
  # y is the mesh points of the matrix
  
  # Fertility matrix
  Fmat<-matrix(0,(n+4),(n+4))
  
  # Fecundity
  
  #Seedlings this year
  Fmat[4,5:(n+4)] <- Prob_Germ_1 * F_z(y, Year_t, ltre_list) + #seedlings this years seed
    Prob_Germ_2 * Prob_Surv_Seed * S1_z(y, Year_t, ltre_list) + #last year seed
    Prob_Germ_2 * (1 - Prob_Germ_1) * Prob_Surv_Seed * Prob_Germ_2 * Prob_Surv_Seed *  S1_z(y, Year_t, ltre_list) #2 year ago seed
  
  #Seedbank with t-1 and t-2 seeds
  Fmat[1,5:(n+4)] <- (1 - Prob_Germ_1) * (1 - Prob_Germ_2) * Prob_Surv_Seed * Prob_Surv_Seed * S1_z(y, Year_t, ltre_list) #3-year old seeds
  Fmat[2,5:(n+4)] <- (1 - Prob_Germ_1) * Prob_Surv_Seed * S1_z(y, Year_t, ltre_list) #2-year old seeds
  Fmat[3,5:(n+4)] <- (1 - Prob_Germ_1 - Prob_Germ_2) * Prob_Surv_Seed * S1_z(y, Year_t, ltre_list) #1-year old seeds
  
  # Growth/survival transition matrix
  Tmat<-matrix(0,(n+4),(n+4))
  
  # probability a seed stays in the seedbank
  Tmat[2,1] <- (1 - Prob_Germ_1) * Prob_Surv_Seed
  Tmat[3,1] <- (1 - Prob_Germ_1) * Prob_Surv_Seed * (1 - Prob_Germ_2) * Prob_Surv_Seed
  Tmat[3,2] <- (1 - Prob_Germ_1) * Prob_Surv_Seed
  
  # probability seed germinates after 2 years
  Tmat[4,1] <- Prob_Germ_2 * Prob_Surv_Seed * (1 - Prob_Germ_2) * Prob_Surv_Seed
  Tmat[4,2] <- Prob_Germ_2 * Prob_Surv_Seed
  
  # probability seed germinates fresh
  Tmat[4,3] <- Prob_Germ_1
  
  # probability a seedling establishes as adult
  Tmat[5:(n+4),4] <- Sur_z1(ltre_list) * Se_z(y, ltre_list)
  
  # survival and growth
  Tmat[5:(n+4),5:(n+4)] <- (outer(y,y,P_z, Year_t, ltre_list)) * h
  
  # Full kernel -- adds two kernels built
  IPMmat <- Fmat + Tmat
  
  out_lamb <- data.frame(lambda = lambda(IPMmat), version = ltre_list$version)
  return(out_lamb)
  
}

# run kernel for eah row of ltre_list
lamb_lists <- lapply(X = ltre_list, FUN = bigmatrix, mat.size = mat.size)

# make into a dataframe
ltre_lambda <- do.call("rbind", lamb_lists)

# add column for sensitivity
ltre_lambda$sensitivity <- NA

# I have to subtract the correct population/year combination
ltre_lambda$Year_t <- sens$Year_t
ltre_lambda$Population <- sens$Population

ltre_lambda_orig <- ltre_lambda$lambda[1:12]

ltre_lambda <- ltre_lambda %>%
  mutate(sensitivity = case_when(Year_t == 1 & Population == 1 ~ (lambda - ltre_lambda_orig[1])/perturbation,
                                 Year_t == 2 & Population == 1 ~ (lambda - ltre_lambda_orig[2])/perturbation,
                                 Year_t == 3 & Population == 1 ~ (lambda - ltre_lambda_orig[3])/perturbation,
                                 Year_t == 1 & Population == 2 ~ (lambda - ltre_lambda_orig[4])/perturbation,
                                 Year_t == 2 & Population == 2 ~ (lambda - ltre_lambda_orig[5])/perturbation,
                                 Year_t == 3 & Population == 2 ~ (lambda - ltre_lambda_orig[6])/perturbation,
                                 Year_t == 1 & Population == 3 ~ (lambda - ltre_lambda_orig[7])/perturbation,
                                 Year_t == 2 & Population == 3 ~ (lambda - ltre_lambda_orig[8])/perturbation,
                                 Year_t == 3 & Population == 3 ~ (lambda - ltre_lambda_orig[9])/perturbation,
                                 Year_t == 1 & Population == 4 ~ (lambda - ltre_lambda_orig[10])/perturbation,
                                 Year_t == 2 & Population == 4 ~ (lambda - ltre_lambda_orig[11])/perturbation,
                                 Year_t == 3 & Population == 4 ~ (lambda - ltre_lambda_orig[12])/perturbation,
                                 TRUE ~ NA_real_)) %>%
  mutate(sensitivity = case_when(version == "original" ~ NA,
                                 TRUE ~ sensitivity)) %>%
  arrange(Year_t, Population, version)

#sensitivity of lambda to each parameter
#Graph <- ltre_lambda %>% filter(version != "original") %>% 
#    ggplot(aes(x = version, y = sensitivity))+
#    geom_point() +
#    facet_grid(Year_t ~ Population) +
#    labs(x = "Vital Rate Coefficient", y = "Sensitivity of lambda to coefficient")+
#    theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Elasticity
original <- sens.params %>%
  pivot_longer(cols = c(1, 4:21), names_to = "perturb", values_to = "coef")%>%
  filter(perturb != "f.slope2") %>%
  filter(perturb != "g.slope2") %>%
  filter(perturb != "se.slope2") %>%
  filter(perturb != "Recruit_sd") %>%
  filter(perturb != "Seeds") %>%
  filter(perturb != "grow.sd") %>%
  arrange(Year_t, Population, perturb)

ltre_lambda <- ltre_lambda %>%
  filter(version != "original")
  
ltre_lambda$coef <- original$coef
ltre_lambda$check <- original$perturb

# Calculate elasticities
ltre_lambda$elasticity <- NA
ltre_lambda <- ltre_lambda %>% 
  mutate(elasticity = sensitivity * as.numeric(coef)/lambda) %>%
  mutate(vr = case_when(version %in% c("f.slope", "fl.int", "se.int", "se.slope") ~ "fecundity",
                        version %in% c("g.int", "g.slope") ~ "growth",
                        version %in% c("s.int", "s.slope") ~ "survival",
                        version == "original" ~ NA_character_,
                        TRUE ~ "discrete")) %>%
  mutate(Population = case_when(Population == 1 ~ "Bayfield",
                                Population == 2 ~ "Fanshawe",
                                Population == 3 ~ "Ilderton",
                                Population == 4 ~ "RBG")) %>%
  # Fix names of things
  mutate(version = case_when(version == "g.int" ~ "Growth Intercept",
                             version == "g.slope" ~ "Growth Slope",
                             version == "se.int" ~ "Seed Production Intercept",
                             version == "se.slope" ~ "Seed Production Slope",
                             version == "fl.int" ~ "Flowering Intercept",
                             version == "f.slope" ~ "Flowering Slope",
                             version == "Recruit_Size" ~ "Recruit Size",
                             version == "s.int" ~ "Survival Intercept",
                             version == "s.slope" ~ "Survival Slope",
                             version == "Prob_Germ_1" ~ "Fresh Germination",
                             version == "Prob_Germ_2" ~ "Seedbank Germination",
                             version == "Prob_Surv_Seed" ~ "Seed Viability",
                             version == "Estab" ~ "Seedling Establishment"))
  
#elasticity of lambda to each parameter

#re-arrange levels for figure
ltre_lambda$version <- factor(ltre_lambda$version, levels = c("Survival Intercept",
                                                                   "Survival Slope",
                                                                   "Growth Intercept",
                                                                   "Growth Slope",
                                                                   "Flowering Intercept",
                                                                   "Flowering Slope",
                                                                   "Seed Production Intercept",
                                                                   "Seed Production Slope",
                                                                   "Seedling Establishment",
                                                                   "Fresh Germination",
                                                                   "Seedbank Germination",
                                                                   "Seed Viability"))

# remove seed bank and seed viability for clarity
ltre_lambda <- ltre_lambda %>%
  filter(version != "Seedbank Germination") %>%
  filter(version != "Seed Viability")

PopFacet <- c("Bayfield" = "(a)", "Fanshawe" = "(b)", "Ilderton"= "(c)", "RBG" = "(d)")

elas.fig <- ltre_lambda %>% 
  filter(version != "original") %>%
  filter(Year_t == 1) %>%
  ggplot(aes(x = version, y = elasticity)) +
  geom_bar(stat = "identity", fill = "darkgray")+
  labs(x = "Model Parameters", y = expression(paste("Elasticity of ", lambda))) +
  facet_wrap(~Population, labeller = as_labeller(PopFacet)) +
  geom_hline(yintercept = 0) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 50, vjust = 1, hjust=1)) +
  theme(strip.background = element_blank()) +
  theme(strip.text = element_text(hjust = 0, size = 12)) +
  #font size
  theme(text = element_text(size = 14))  

elas.fig

# ggsave(plot = elas.fig, dpi = 600, width = 12, height = 6,
#        filename = "Elasticity_ParamsYr1.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = elas.fig, dpi = 600, width = 12, height = 6,
#        filename = "Elasticity_ParamsYr1.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")

### make it with multiple panels
p <- ltre_lambda %>% 
    filter(version != "original") %>%
    filter(Year_t == 1) %>%
    filter(Population == "Bayfield") %>%
    ggplot(aes(x = version, y = elasticity)) +
    geom_bar(stat = "identity", fill = "#D55E00")+
    labs(x = "Model Parameters", y = expression(paste("Elasticity of ", lambda))) +
  scale_y_continuous(limits = c(-2.1, 0.7)) +
    geom_hline(yintercept = 0) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12)) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.0)))

p2 <- ltre_lambda %>% 
    filter(version != "original") %>%
    filter(Year_t == 1) %>%
    filter(Population == "Fanshawe") %>%
    ggplot(aes(x = version, y = elasticity)) +
    geom_bar(stat = "identity", fill = "#0072B2")+
    labs(x = "Model Parameters", y = expression(paste("Elasticity of ", lambda))) +
    theme(axis.text.x = element_text(angle = 0, hjust = 1)) +
    scale_y_continuous(limits = c(-2.1, 0.7)) +
    geom_hline(yintercept = 0) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.title = element_blank(),
        axis.text = element_blank()) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.0)))

p3 <- ltre_lambda %>% 
    filter(version != "original") %>%
    filter(Year_t == 1) %>%
    filter(Population == "Ilderton") %>%
    ggplot(aes(x = version, y = elasticity)) +
    geom_bar(stat = "identity", fill = "#F0E442")+
    labs(x = "Model Parameters", y = expression(paste("Elasticity of ", lambda))) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    theme(axis.title = element_text(size = 14)) +
      scale_y_continuous(limits = c(-1.8, 2)) +
    geom_hline(yintercept = 0) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.title = element_text(size = 14),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 14),
        axis.text.y = element_text(size = 12)) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.0)))

p4 <- ltre_lambda %>% 
    filter(version != "original") %>%
    filter(Year_t == 1) %>%
    filter(Population == "RBG") %>%
    ggplot(aes(x = version, y = elasticity)) +
    geom_bar(stat = "identity", fill = "#009E73")+
    labs(x = "Model Parameters", y = expression(paste("Elasticity of ", lambda))) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_y_continuous(limits = c(-1.8, 2)) +
    geom_hline(yintercept = 0) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 14),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.0)))

Graphs <- ggarrange(p,  p2,  p3,  p4,
                      labels = c("(a) Bayfield", "(b) Fanshawe", "(c) Ilderton", "(d) RBG"),
                      font.label = list(size = 14), 
                      hjust = c(-0.7, -0.2, -0.7, -0.3),#c(-3.5,-0.7,-3.5, -0.7),
                    heights = c(1, 1.5),
                      ncol = 2, nrow = 2#, align = "hv"
                    )
Graphs

# ggsave(plot = Graphs, dpi = 600, width = 12, height = 10,
#        filename = "Elasticity_ParamsYr1.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = Graphs, dpi = 600, width = 12, height = 10,
#        filename = "Elasticity_ParamsYr1.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
```


Make supplemental figure with all the elasticities for all years
```{r}
# change facet labels
PopFacet <- c("Bayfield" = "Population 1", "Fanshawe" = "Population 3", "Ilderton"= "Population 4", "RBG" = "Population 6",
              "1" = "2022", "2" = "2023", "3" = "2024")
# change facet labels
PopFacet <- c("Bayfield" = "Bayfield", "Fanshawe" = "Fanshawe", "Ilderton"= "Ilderton", "RBG" = "RBG",
              "1" = "2022", "2" = "2023", "3" = "2024")

# Make full figure
elas.fig.all <- ltre_lambda %>% 
  filter(version != "original") %>%
  #filter(Year_t == 1) %>%
  ggplot(aes(x = version, y = elasticity, fill = Population)) +
  geom_bar(stat = "identity")+
  labs(x = "Model Parameters", y = expression(paste("Elasticity of ", lambda))) +
  facet_grid(Year_t ~ Population, labeller = as_labeller(PopFacet)) +
  scale_fill_manual(values = c("#D55E00", "#0072B2", "#F0E442", "#009E73")) +
  geom_hline(yintercept = 0) +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 50, vjust = 1, hjust=1)) +
  theme(strip.background = element_blank()) +
  theme(strip.text = element_text(hjust = 0, size = 12)) +
  #font size
  theme(text = element_text(size = 14)) 

elas.fig.all

# ggsave(plot = elas.fig.all, dpi = 600, width = 12, height = 6,
#        filename = "Elasticity_Params_AllYr.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = elas.fig.all, dpi = 600, width = 12, height = 6,
#        filename = "Elasticity_Params_AllYr.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
```

# Simulations

code from: https://github.com/jloesberg/Primula/blob/main/scripts/primula_IPM.rmd

These simulations are bootstrapped, so I need to reload the bootstrapped
 model (params_list, functions) from above, rather than the deterministic
 versions.

## Seedling Survival

```{r}
# Probability a seedling survives
Estab <- est.intercepts

bigmatrix <- function(params_list, mat.size, Estab){
  n <- mat.size #matrix size
  L <- min.size #lower integration limit
  U <- max.size #upper integration limit
 
  h <- (U - L)/n #Bin size
  b <- L + c(0:n) * h #Lower boundaries 
  y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
  # y is the mesh points of the matrix
  
  # Fertility matrix
  Fmat<-matrix(0,(n+3),(n+3))
  
  # Fecundity
  
  #Seedlings this year
  Fmat[3,4:(n+3)] <- Prob_Germ_1 * F_z(y, Year_t, params_list) + #seedlings this years seed
    Prob_Germ_2 * Prob_Surv_Seed * S1_z(y, Year_t, params_list) + #last year seed
    Prob_Germ_2 * (1 - Prob_Germ_1) * Prob_Surv_Seed * Prob_Germ_2 * Prob_Surv_Seed *  S1_z(y, Year_t, params_list) #2 year ago seed
  
  #Seedbank with t-1 and t-2 seeds
  Fmat[1,4:(n+3)] <- (1 - Prob_Germ_1) * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #2-year old seeds
  Fmat[2,4:(n+3)] <- (1 - Prob_Germ_1 - Prob_Germ_2) * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #1-year old seeds
  
  # Growth/survival transition matrix
  Tmat<-matrix(0,(n+3),(n+3))
  
  # probability a seed stays in the seedbank
  Tmat[2,1] <- 1 - Prob_Germ_1 * Prob_Surv_Seed
  
  # probability seed germinates after 2 years
  Tmat[3,1] <- Prob_Germ_2 * Prob_Surv_Seed
  
  # probability seed germinates fresh
  Tmat[3,2] <- Prob_Germ_1
  
  # probability a seedling establishes as adult
  Tmat[4:(n+3),3] <- Estab * Se_z(y, params_list)
  
  # survival and growth
  Tmat[4:(n+3),4:(n+3)] <- (outer(y,y,P_z, Year_t, params_list)) * h #removed transpose of outer()
  
  # Full kernel -- adds two kernels built
  IPMmat <- Fmat + Tmat
  
  #out_matrix <- list(IPMmat = IPMmat, Fmat = Fmat, Tmat = Tmat, meshpts = y)
  
  out_lamb <- data.frame(lambda = lambda(IPMmat))
  #return(out_matrix) #outputs matrices
  return(out_lamb) #output lambda
}

# range of seedling survival, 200 samples
# from 0 to 0.5
prob_values =  seq(0.000, 0.5, .01)

# Create an empty list to store the results
results_list <- list()

# Loop through each value
for (prob_value in prob_values) {
  # Call the bigmatrix function with the current prob_value
  
  result <- lapply(X = params_list, FUN = bigmatrix, mat.size = mat.size, Estab = prob_value)
  
  # Add the result to the list with an identifier
  results_list[[as.character(prob_value)]] <- result
}

# saveRDS(results_list, file = "~/Research/Data/IPM_Formatted_STYL/Output/SeedlingSurvSim.rds")
results_list <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/SeedlingSurvSim.rds")


# first make an empty list
datalist <- vector(mode = "list", length = 51)

# make a loop to extract values
for(i in 1:51){
  SeedSurv_lambda_a <- do.call(rbind, results_list[[i]])
  SeedSurv_lambda_a <- tibble::rownames_to_column(SeedSurv_lambda_a, "Prob_Surv_Seed")
  SeedSurv_lambda_a$Prob_Surv_Seed <- prob_values[i]
  SeedSurv_lambda_a$Pop <- c(rep(1, 3000), rep(2, 3000), rep(3, 3000), rep(4, 3000))
  SeedSurv_lambda_a$Year <- rep(c(1, 2, 3), times = 4000)
  SeedSurv_lambda_a <- SeedSurv_lambda_a %>%
    # use the first two groups to split by population and year
    group_by(#Pop, Year,###
             Prob_Surv_Seed) %>%
    summarise(mean = mean(lambda),
              LCI = quantile(lambda, probs = 0.025),
              UCI = quantile(lambda, probs = 0.975))
  datalist[[i]] <- SeedSurv_lambda_a
}
# Extract dataframes and rbind them
SeedSurv_lambda <- bind_rows(datalist)


# Fix pop names and years if needed
SeedSurv_lambda <- SeedSurv_lambda %>%
    mutate(Pop = case_when(Pop == 1 ~ "Bayfield",
                                Pop == 2 ~ "Fanshawe",
                                Pop == 3 ~ "Ilderton",
                                Pop == 4 ~ "RBG"),
         Year = case_when(Year == 1 ~ 2022,
                            Year == 2 ~ 2023,
                            Year == 3 ~ 2024))

Obs_lambda <- SeedSurv_lambda %>%
  filter(Prob_Surv_Seed > 0.18) %>%
  filter(Prob_Surv_Seed < 0.20)

# Combined all  -- have to comment out group_by in all the SeedSurv dataframes
p <- ggplot(SeedSurv_lambda %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), aes(x = Prob_Surv_Seed, y = mean)) +
  geom_line(color = "black", linewidth = 1) +
  geom_ribbon(SeedSurv_lambda %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), mapping = aes(ymin = LCI, ymax = UCI), alpha = 0.3) +
  # observed value point
  geom_point(Obs_lambda %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), mapping = aes(x = Prob_Surv_Seed, y = mean), color = "darkblue", size = 3) +
  # lambda of 1 line
  geom_hline(yintercept = 1, color = "red", linetype = 2) +
  labs(x = "Seedling Establishment (%)", y = expression(paste("Growth rate (", lambda, ")"))) +
  #facet_grid(Pop ~ Year, scales = "free") +
  theme_classic() +
  theme(text = element_text(size = 16))
p
# ggsave(plot = p, dpi = 600, width = 12, height = 8, # was 12w x 6 h
#        filename = "Seedling_Surv_Sim.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = p, dpi = 600, width = 12, height = 8, # was 12w x 6 h
#        filename = "Seedling_Surv_Sim.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")

# Make separate panels for each year/pop combination for the supplemental
SeedSurv_lambda2 <- SeedSurv_lambda %>%
  mutate(Pop = case_when(Pop == "Bayfield" ~ "1",
                         Pop == "Fanshawe" ~ "3",
                         Pop == "Ilderton" ~ "4",
                         Pop == "RBG" ~ "6",
                         TRUE ~ NA_character_))

pAll <- ggplot(SeedSurv_lambda %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), aes(x = Prob_Surv_Seed, y = mean)) +
  geom_line(color = "black", linewidth = 1) +
  geom_ribbon(SeedSurv_lambda2 %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), mapping = aes(ymin = LCI, ymax = UCI), alpha = 0.3) +
  # observed value point
  geom_point(Obs_lambda %>%
              # mutate(Pop = case_when(Pop == "Bayfield" ~ "1",
              #                        Pop == "Fanshawe" ~ "3",
              #                        Pop == "Ilderton" ~ "4",
              #                        Pop == "RBG" ~ "6",
              #                        TRUE ~ NA_character_)) %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), mapping = aes(x = Prob_Surv_Seed, y = mean), color = "darkblue", size = 3) +
  # lambda of 1 line
  geom_hline(yintercept = 1, color = "red", linetype = 2) +
  labs(x = "Seedling Establishment (%)", y = expression(paste("Growth rate (", lambda, ")"))) +
  facet_grid(Pop ~ Year, scales = "free_y") +
  theme_classic() +
  theme(text = element_text(size = 16)) +
  theme(strip.background = element_blank())

pAll

# ggsave(plot = pAll, dpi = 600, width = 10, height = 8, # was 12w x 6 h
#        filename = "Seedling_Surv_Sim_Split.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = pAll, dpi = 600, width = 10, height = 8, # was 12w x 6 h
#        filename = "Seedling_Surv_Sim_Split.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")

```

## Seedling Germination
```{r}
# Probability a seedling survives
Prob_Germ_1

Prob_Germ_1_Orig <- Prob_Germ_1

bigmatrix <- function(params_list, mat.size, Prob_Germ_1){
  n <- mat.size #matrix size
  L <- min.size #lower integration limit
  U <- max.size #upper integration limit
 
  h <- (U - L)/n #Bin size
  b <- L + c(0:n) * h #Lower boundaries 
  y <- 0.5 * (b[1:n] + b[2:(n+1)])  #Bin midpoints
  # y is the mesh points of the matrix
  
  # Fertility matrix
  Fmat<-matrix(0,(n+3),(n+3))
  
  # Fecundity
  
  #Seedlings this year
  Fmat[3,4:(n+3)] <- Prob_Germ_1 * F_z(y, Year_t, params_list) + #seedlings this years seed
    Prob_Germ_2 * Prob_Surv_Seed * S1_z(y, Year_t, params_list) + #last year seed
    Prob_Germ_2 * (1 - Prob_Germ_1_Orig) * Prob_Surv_Seed * Prob_Germ_2 * Prob_Surv_Seed *  S1_z(y, Year_t, params_list) #2 year ago seed
  
  #Seedbank with t-1 and t-2 seeds
  Fmat[1,4:(n+3)] <- (1 - Prob_Germ_1_Orig) * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #2-year old seeds
  Fmat[2,4:(n+3)] <- (1 - Prob_Germ_1_Orig - Prob_Germ_2) * Prob_Surv_Seed * S1_z(y, Year_t, params_list) #1-year old seeds
  
  # Growth/survival transition matrix
  Tmat<-matrix(0,(n+3),(n+3))
  
  # probability a seed stays in the seedbank
  Tmat[2,1] <- 1 - Prob_Germ_1 * Prob_Surv_Seed
  
  # probability seed germinates after 2 years
  Tmat[3,1] <- Prob_Germ_2 * Prob_Surv_Seed
  
  # probability seed germinates fresh
  Tmat[3,2] <- Prob_Germ_1
  
  # probability a seedling establishes as adult
  Tmat[4:(n+3),3] <- Sur_z1(params_list) * Se_z(y, params_list)
  
  # survival and growth
  Tmat[4:(n+3),4:(n+3)] <- (outer(y,y,P_z, Year_t, params_list)) * h #removed transpose of outer()
  
  # Full kernel -- adds two kernels built
  IPMmat <- Fmat + Tmat
  
  #out_matrix <- list(IPMmat = IPMmat, Fmat = Fmat, Tmat = Tmat, meshpts = y)
  
  out_lamb <- data.frame(lambda = lambda(IPMmat))
  #return(out_matrix) #outputs matrices
  return(out_lamb) #output lambda
}

# range of seedling germination survival
# from 0 to 0.5
prob_values =  seq(0.000, 0.5, 0.01)

# Create an empty list to store the results
results_list <- list()

# Loop through each value
for (prob_value in prob_values) {
  
  # Call the bigmatrix function with the current prob_value
  result <- lapply(X = params_list, FUN = bigmatrix, mat.size = mat.size, Prob_Germ_1 = prob_value)
  
  # Add the result to the list with an identifier
  results_list[[as.character(prob_value)]] <- result
}

# saveRDS(results_list, file = "~/Research/Data/IPM_Formatted_STYL/Output/SeedlingGermSim.rds")

 results_list <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/SeedlingGermSim.rds")

 #add in the values used:
 
 # first make an empty list
datalist <- vector(mode = "list", length = 51)

# make a loop to extract values
for(i in 1:51){
  SeedSurv_lambda_a <- do.call(rbind, results_list[[i]])
  SeedSurv_lambda_a <- tibble::rownames_to_column(SeedSurv_lambda_a, "Prob_Surv_Seed")
  SeedSurv_lambda_a$Prob_Surv_Seed <- prob_values[i]
  SeedSurv_lambda_a$Pop <- c(rep(1, 3000), rep(2, 3000), rep(3, 3000), rep(4, 3000))
  SeedSurv_lambda_a$Year <- rep(c(1, 2, 3), times = 4000)
  SeedSurv_lambda_a <- SeedSurv_lambda_a %>%
    # use the first two groups to split by population and year
    group_by(#Pop, Year,###
             Prob_Surv_Seed) %>%
    summarise(mean = mean(lambda),
              LCI = quantile(lambda, probs = 0.025),
              UCI = quantile(lambda, probs = 0.975))
  datalist[[i]] <- SeedSurv_lambda_a
}
# Extract dataframes and rbind them
SeedSurv_lambda <- bind_rows(datalist)

# Fix pop names and years if needed
#SeedSurv_lambda <- SeedSurv_lambda %>%
#    mutate(Pop = case_when(Pop == 1 ~ "Bayfield",
#                                Pop == 2 ~ "Fanshawe",
#                                Pop == 3 ~ "Ilderton",
#                                Pop == 4 ~ "RBG"),
#         Year = case_when(Year == 1 ~ 2022,
#                            Year == 2 ~ 2023,
#                            Year == 3 ~ 2024))

Obs_lambda <- SeedSurv_lambda %>%
  filter(Prob_Surv_Seed > 0.04) %>%
  filter(Prob_Surv_Seed < 0.06) %>%
  mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100)


###########################################
########################################
p2 <- ggplot(SeedSurv_lambda %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), aes(x = Prob_Surv_Seed, y = mean)) +
  geom_line(color = "black", linewidth = 1) +
  geom_ribbon(SeedSurv_lambda %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), mapping = aes(ymin = LCI, ymax = UCI), alpha = 0.3) +
  # observed value point
  geom_point(Obs_lambda, mapping = aes(x = Prob_Surv_Seed, y = mean), color = "darkblue", size = 3) +
  # lambda of 1 line
  geom_hline(yintercept = 1, color = "red", linetype = 2) +
  labs(x = "Seedling Emergence (%)", y = expression(paste("Growth rate (", lambda, ")"))) +
  #facet_grid(Pop ~ Year, scales = "free") +
  theme_classic() +
  theme(text = element_text(size = 16))

p2

# ggsave(plot = p2, dpi = 600, width = 12, height = 8, # was 12w x 6 h
#        filename = "Seedling_Germ_Sim.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = p2, dpi = 600, width = 12, height = 8, # was 12w x 6 h
#        filename = "Seedling_Germ_Sim.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")

# Make a figure separating the year/pop combinations for the supplemental
# Just requires adding to the group_by for each SeedSurv dataframe

SeedSurv_lambda2 <- SeedSurv_lambda %>%
  #mutate(Pop = case_when(Pop == "Bayfield" ~ "1",
  #                       Pop == "Fanshawe" ~ "3",
  #                       Pop == "Ilderton" ~ "4",
  #                       Pop == "RBG" ~ "6",
  #                       TRUE ~ NA_character_)) %>%
  mutate(Pop = case_when(Pop == "1" ~ "Bayfield",
                         Pop == "2" ~ "Fanshawe",
                         Pop == "3" ~ "Ilderton",
                         Pop == "4" ~ "RBG",
                         TRUE ~ NA_character_),
         Year = case_when(Year == "1" ~ "2022",
                          Year == "2" ~ "2023",
                          Year == "3" ~ "2024"))


pAll <- ggplot(SeedSurv_lambda2 %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), aes(x = Prob_Surv_Seed, y = mean)) +
  geom_line(color = "black", linewidth = 1) +
  geom_ribbon(SeedSurv_lambda2 %>%
              # need to make a percent!
              mutate(Prob_Surv_Seed = Prob_Surv_Seed * 100), mapping = aes(ymin = LCI, ymax = UCI), alpha = 0.3) +
  # observed value point
  geom_point(Obs_lambda %>%
               mutate(Pop = case_when(Pop == "1" ~ "Bayfield",
                         Pop == "2" ~ "Fanshawe",
                         Pop == "3" ~ "Ilderton",
                         Pop == "4" ~ "RBG",
                         TRUE ~ NA_character_),
         Year = case_when(Year == "1" ~ "2022",
                          Year == "2" ~ "2023",
                          Year == "3" ~ "2024"))
             , mapping = aes(x = Prob_Surv_Seed, y = mean), color = "darkblue", size = 3) +
  # lambda of 1 line
  geom_hline(yintercept = 1, color = "red", linetype = 2) +
  labs(x = "Seedling Emergence (%)", y = expression(paste("Growth rate (", lambda, ")"))) +
  facet_grid(Pop ~ Year, scales = "free_y") +
  theme_classic() +
  theme(text = element_text(size = 16)) +
  theme(strip.background = element_blank())

pAll

# ggsave(plot = pAll, dpi = 600, width = 10, height = 8, # was 12w x 6 h
#        filename = "Seedling_Germ_Sim_Split.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = pAll, dpi = 600, width = 10, height = 8, # was 12w x 6 h
#        filename = "Seedling_Germ_Sim_Split.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
```

Make a figure with the two simulations
This is why they are p and p2
```{r}
Graphs <- ggarrange(p, p2,
                    labels = c("(a)", "(b)"),
                    font.label = list(size = 14), 
                    hjust = c(-4.0, -4.0),
                    ncol = 1, nrow = 2, align = "v")
Graphs

# ggsave(plot = Graphs, dpi = 600, width = 12, height = 8, # was 12w x 6 h
#        filename = "Combined_Sims.png",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
# ggsave(plot = Graphs, dpi = 600, width = 12, height = 8, # was 12w x 6 h
#        filename = "Combined_Sims.pdf",
#        path = "C:/Users/awied/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs")
```

