---
title: "Wood-poppy Integrated Population Model"
author: "A"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

Construct an integrated population model (IPM) for the wood-poppy,
using multiple sets of data

Load data for the model. This includes the capture histories for each
 plant, the wood-poppy size data for each year, population counts,
 and the number of untagged plants counted at each population
```{r message=FALSE, warning=FALSE}
#Load libraries
library(tidyverse)
library(dplyr)
library(stringr)
library(IPMbook)
library(jagsUI)
library(ggh4x) # for plots

#Load data
capt_hist <- read.csv("~/Research/Data/Github Code/Stylophorum-IPMs/MarkRecapture_Multistate.csv", colClasses='character')
# Horizontal format of 
Wood_Poppy_Data <- read.csv("~/Research/Data/Github Code/Stylophorum-IPMs/Wood_Poppy_All.csv")
#Vertical merge of tagged plants
Wood_Poppy_Vertical <- read.csv("~/Research/Data/Github Code/Stylophorum-IPMs/Wood_Poppy_Vertical.csv")
#population counts
Counts <- read.csv("~/Research/Data/Github Code/Stylophorum-IPMs/Count_Data.csv")
#untagged but counted plants
Untagged <- read.csv("~/Research/Data/Github Code/Stylophorum-IPMs/Untagged_Data.csv")
# Data on individual plants from 2020-2021
WSP_Data <- read.csv("~/Research/Data/Github Code/Stylophorum-IPMs/WSP_Data_Clean.csv")

```

Number of Populations
```{r echo=FALSE}
popn <- Wood_Poppy_Data %>%
  distinct(Population) %>%
  arrange(Population) %>%
  mutate(PopN = 1:length(unique(Wood_Poppy_Data$Population)))

npop <- nrow(popn)
npop
```

# Format Data

Restructure the data to fit the format for the models
Data need to be formatted so that populations are rows and years are
 columns, always beginning with the first year data were collected (2020),
 even if data were not collected at a particular site that year. The
 populations also need to be in alphabetical order, so that the rows
 match up in each dataset.
 
```{r}
################################################################################
########## Capture Histories ###################################################
################################################################################


################################################################################
########## Population Counts ###################################################
################################################################################
#For seed data below, need to know the proportion of plants that
# were tagged at Ilderton
Ilderton_Prop <- Wood_Poppy_Vertical %>%
  filter(Population == "Ilderton") %>%
  group_by(Year) %>%
  summarise(n = n()) %>%
  full_join(Counts %>%
              filter(Population == "Ilderton")) %>%
  mutate(Prop_of_Tot = n/Count) %>%
  ungroup() %>%
  summarise(Prop_of_Tot = mean(Prop_of_Tot, na.rm = T)) %>%
  pull(Prop_of_Tot)



Counts[nrow(Counts) + 1, ] <- c("Stylophorum", "Meadowlily", 2023, NA_integer_)
Counts[nrow(Counts) + 1, ] <- c("Stylophorum", "Meadowlily", 2024, NA_integer_)
Counts[nrow(Counts) + 1, ] <- c("Stylophorum", "Ilderton", 2022, NA_integer_)
Counts[nrow(Counts) + 1, ] <- c("Stylophorum", "RBG", 2020, NA_integer_)
Counts[nrow(Counts) + 1, ] <- c("Stylophorum", "RBG", 2022, NA_integer_)
Counts[nrow(Counts) + 1, ] <- c("Stylophorum", "RBG", 2023, NA_integer_)
Counts[nrow(Counts) + 1, ] <- c("Stylophorum", "RBG", 2024, NA_integer_)
Counts <- Counts %>%
    left_join(popn, by = "Population") %>%
  filter(Species == "Stylophorum") %>%
  mutate_at(c('Year', 'Count'), as.numeric) %>%
  arrange(Population, Year) %>%
  rename(Counts = Count) %>%
  dplyr::select(PopN, Year, Counts)

#fix format
Counts <- Counts %>%
  pivot_wider(names_from = Year, values_from = Counts) %>%
  column_to_rownames(var = "PopN")
```

Remove seedlings from count data
```{r}
Counts2 <- Wood_Poppy_Vertical %>%
  group_by(Population, Year) %>%
  filter(Height_cm > 9) %>%
  summarise(Plants = n())

Counts3 <- WSP_Data %>%
  mutate(Year = year(ymd(Date))) %>%
  group_by(Population, Year) %>%
  summarise(Small = sum(Less_15_cm, na.rm = T),
            Big = sum(More_15_cm, na.rm = T)) %>%
  filter(!is.na(Small)) %>%
  rowwise() %>%
  mutate(Plants = sum(c_across(Small:Big))) %>%
  dplyr::select(Population, Year, Plants)


Counts4 <- Untagged %>%
  group_by(Population, Year) %>%
  summarise(Small = sum(Less_15_cm),
            Big = sum(More_15_cm)) %>%
  filter(!is.na(Small)) %>%
  rowwise() %>%
  mutate(Plants2 = sum(c_across(Small:Big))) %>%
  dplyr::select(Population, Year, Plants2) %>%
  ungroup() %>%
  full_join(Counts2) %>%
  group_by(Population, Year) %>%
  summarise(Plants2 = sum(Plants2),
            Plants = sum(Plants)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(Plants3 = sum(c_across(Plants2:Plants), na.rm = T)) %>%
  dplyr::select(Population, Year, Plants = Plants3)

Counts3 <- rbind(Counts4, Counts3)

Counts3 <- Counts3 %>%
  group_by(Population, Year) %>%
  summarise(Plants = sum(Plants)) %>%
  ungroup() %>%
  pivot_wider(names_from = Year, values_from = Plants) %>%
  dplyr::select(Population, `2020`, `2021`, `2022`, `2023`, `2024`, `2025`) %>%
  dplyr::select(-Population)
  
Counts3

Counts3[2,5] <- 0 # only seedlings at Denfield
Counts3[2,6] <- 0 # only seedlings at Denfield
Counts3[6,3] <- NA
Counts3[6,4] <- NA
Counts3[6,5] <- NA
Counts3[6,2] <- 542

# Fix Ilderton
Counts3[4,3] <- NA
Counts3[4,4] <- 1145
Counts3[4,5] <- 927
Counts3[4,6] <- 1174
```

Combine all needed data
```{r}

# Different start years for recapture at different sites
# WSP did not tag at Ilderton or RBG, so no capture data
# until 2022
start <- c(1, 1, 1, 3, 1, 3)
end <- c(6, 6, 6, 6, 3, 6)
# Set the start and end years for the population counts
# RBG was only counted in 2021 and 2025, and no data on Meadowlily
# after 2022
start_c <- c(1, 1, 1, 1, 1, 2)
end_c <- c(6, 6, 6, 6, 3, 6)

#set initial population sizes
pinit <- dUnif(lower = rep(1, 6), upper = apply(Counts, 1, max, na.rm = T))

# Set initial seed number sizes--this is the max number of seeds
# produced in a year at each population (see the priors section)
#pinitS <- dUnif(lower = c(1,1,1,1,1,1), upper = c( 11696,2079, 136063,2079,37424,19956))
pinitS <- c(11696,500, 136063,32079,37424,19956)

# Gather up all the data into a list for JAGS
# Be sure that the m-arrays are formatted properly,
# and that the number of occasions dimension is correct
# n.occ should be 5 (for now)
jags.data <- list(n.occ = 5, # 5 capture occasions
                  Counts = Counts3,
                  npop = npop,
                  start = start,
                  end = end,
                  start_c = start_c,
                  end_c = end_c,
                  pinit = pinit,
                  pinitS = pinitS#,
                  #K = K
                  )

# Save jags.data as an RDS object so I don't have to run everything all the time
# saveRDS(jags.data, file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_25.rds")
```

## Multistate Mark Recapture
```{r}
jags.data <- readRDS("~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_25.rds")

################################################################################
########## Capture Histories ###################################################
################################################################################
# First up, stage of each plant at the first time it was seen
Age <- Wood_Poppy_Vertical %>%
  # Add population numbers (rather than names)
  left_join(popn, by = "Population") %>%
  dplyr::select(PopN, Individual, Year, Height_Class) %>%
  mutate(Height_Class = case_when(Height_Class == "0" ~ NA_integer_,
                                  TRUE ~ Height_Class)) %>%
  group_by(PopN, Individual) %>%
  mutate(first = dplyr::first(na.omit(Height_Class))) %>%
  distinct(Individual, .keep_all = T) %>%
  filter(!is.na(first)) %>% #####
ungroup() %>%
  arrange(PopN, Individual) %>%
  dplyr::select(-Height_Class)

Ages <- Age %>%
  pull(first)

length(Ages)

# Capture Histories for each plant
Capt_hist2 <- capt_hist %>%
  left_join(popn, by = "Population") %>%
  right_join(Age, by = c("PopN", "Individual")) %>% #keep only ones we have ages for
  rename("Year_1" = "X2020", "Year_2" = "X2021", "Year_3" = "X2022",
         "Year_4" = "X2023", "Year_5" = "X2024", "Year_6" = "X2025") %>%
  mutate_at(c("Year_1", "Year_2", "Year_3", "Year_4", "Year_5", "Year_6"), as.numeric) %>%
  #Meadowlily wasn't surveyed, so there is no data--note that this means
  # the new tags in 2022 simply 'vanish' in the model
  mutate(Year_4 = case_when(Population == "Meadowlily" ~ NA_real_,
                            TRUE ~ Year_4)) %>%
  mutate(Year_5 = case_when(Population == "Meadowlily" ~ NA_real_,
                            TRUE ~ Year_5)) %>%
  mutate(Year_6 = case_when(Population == "Meadowlily" ~ NA_real_,
                            TRUE ~ Year_6)) %>%
  arrange(Population, Individual) %>%
  dplyr::select(-Population, -Individual, -Year, -first, -PopN) 

head(Capt_hist2)
str(Capt_hist2)

nrow(Capt_hist2)

# Format population numbers the same as the capture histories
Population <- Wood_Poppy_Vertical %>%
  #anti_join(Sizes, by = c("Population", "Individual")) %>%
  left_join(popn, by = "Population") %>%
  dplyr::select(PopN, Individual, Year, Height_Class) %>%
  mutate(Height_Class = case_when(Height_Class == "0" ~ NA_integer_,
                                  TRUE ~ Height_Class)) %>%
  group_by(PopN, Individual) %>%
  mutate(first = dplyr::first(na.omit(Height_Class))) %>%
  distinct(Individual, .keep_all = T) %>%
  filter(!is.na(first)) %>% #####
ungroup() %>%
  arrange(PopN, Individual) %>%
  pull(PopN)
################################################################################
########## M-Array #############################################################
################################################################################

# Capture Histories for each plant
Capt_hist2 <- capt_hist %>%
  left_join(popn, by = "Population") %>%
  right_join(Age, by = c("PopN", "Individual")) %>% #keep only ones we have ages for
  rename("Year_1" = "X2020", "Year_2" = "X2021", "Year_3" = "X2022",
         "Year_4" = "X2023", "Year_5" = "X2024", "Year_6" = "X2025") %>%
  mutate_at(c("Year_1", "Year_2", "Year_3", "Year_4", "Year_5", "Year_6"), as.numeric) %>%
  #Meadowlily wasn't surveyed, so there is no data--note that this means
  # the new tags in 2022 simply 'vanish' in the model
  mutate(Year_4 = case_when(Population == "Meadowlily" ~ NA_real_,
                            TRUE ~ Year_4)) %>%
  mutate(Year_5 = case_when(Population == "Meadowlily" ~ NA_real_,
                            TRUE ~ Year_5)) %>%
  mutate(Year_6 = case_when(Population == "Meadowlily" ~ NA_real_,
                            TRUE ~ Year_6)) %>%
  arrange(Population, Individual)

##### Bayfield #############################################
Bmarr <- Capt_hist2 %>%
  filter(Population == "Bayfield") %>%
  dplyr::select(-Population, -Individual, -Year, -first, -PopN)
Bmarr <- as.matrix(Bmarr)

marr <- marray(Bmarr)
# Bundle data
# Calculate the number of states
ns <- length(unique(as.numeric(Bmarr))) - 1

rel  <- rowSums(marr)
zero <- matrix(0, ns, ns)
ones <- diag(ns)

# Subset jags.data (loaded above) for just population 1
jags.data$n.occ <- 6
jags.data$Counts <- as.vector(t(jags.data$Counts[1,]))
jags.data$npop <- 1
jags.data$start <- jags.data$start[1]
jags.data$start_c <- jags.data$start_c[1]
jags.data$end <- jags.data$end[1]
jags.data$end_c <- jags.data$end_c[1]
jags.data$pinit <- jags.data$pinit[1,]
jags.data$pinitS <- jags.data$pinitS[1]
jags.data$K <- 0

jags.data2 <- jags.data

jags.data2$marr <- marr
jags.data2$ns <- ns
jags.data2$rel <- rel
jags.data2$zero <- zero
jags.data2$ones <- ones

jags.data2$nyears <-jags.data2$n.occ

# saveRDS(jags.data2, file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Bayfield.rds")

##### Denfield #############################################
Dmarr <- Capt_hist2 %>%
  filter(Population == "Denfield") %>%
  dplyr::select(-Population, -Individual, -Year, -first, -PopN)
Dmarr <- as.matrix(Dmarr)

marr <- marray(Dmarr)
# Bundle data
# Calculate the number of states
ns <- length(unique(as.numeric(Dmarr))) - 1

rel  <- rowSums(marr)
zero <- matrix(0, ns, ns)
ones <- diag(ns)

jags.data$n.occ <- 6
jags.data$Counts <- as.vector(t(jags.data$Counts[2,]))
jags.data$npop <- 2
jags.data$start <- jags.data$start[2]
jags.data$start_c <- jags.data$start_c[2]
jags.data$end <- jags.data$end[2]
jags.data$end_c <- jags.data$end_c[2]
jags.data$pinit <- jags.data$pinit[2,]
jags.data$pinitS <- jags.data$pinitS[2]
jags.data$K <- 0

jags.data2 <- jags.data

jags.data2$marr <- marr
jags.data2$ns <- ns
jags.data2$rel <- rel
jags.data2$zero <- zero
jags.data2$ones <- ones

jags.data2$nyears <-jags.data2$n.occ

# saveRDS(jags.data2, file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Denfield.rds")

##### Fanshawe #############################################
Fmarr <- Capt_hist2 %>%
  filter(Population == "Fanshawe") %>%
  dplyr::select(-Population, -Individual, -Year, -first, -PopN)
Fmarr <- as.matrix(Fmarr)

marr <- marray(Fmarr)
# Bundle data
# Calculate the number of states
ns <- length(unique(as.numeric(Fmarr))) - 1

rel  <- rowSums(marr)
zero <- matrix(0, ns, ns)
ones <- diag(ns)

jags.data$n.occ <- 6
jags.data$Counts <- as.vector(t(jags.data$Counts[3,]))
jags.data$npop <- 3
jags.data$start <- jags.data$start[3]
jags.data$start_c <- jags.data$start_c[3]
jags.data$end <- jags.data$end[3]
jags.data$end_c <- jags.data$end_c[3]
jags.data$pinit <- jags.data$pinit[3,]
jags.data$pinitS <- jags.data$pinitS[3]
jags.data$K <- 0

jags.data2 <- jags.data

jags.data2$marr <- marr
jags.data2$ns <- ns
jags.data2$rel <- rel
jags.data2$zero <- zero
jags.data2$ones <- ones

jags.data2$nyears <-jags.data2$n.occ
# saveRDS(jags.data2, file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Fanshawe.rds")

##### Ilderton #############################################
Imarr <- Capt_hist2 %>%
  filter(Population == "Ilderton") %>%
  dplyr::select(-Population, -Individual, -Year, -first, -PopN)
Imarr <- as.matrix(Imarr)

marr <- marray(Imarr)
# Bundle data
# Calculate the number of states
ns <- length(unique(as.numeric(Imarr))) - 1

rel  <- rowSums(marr)
zero <- matrix(0, ns, ns)
ones <- diag(ns)

jags.data$n.occ <- 6
jags.data$Counts <- as.vector(t(jags.data$Counts[4,]))
jags.data$npop <- 4
jags.data$start <- jags.data$start[4]
jags.data$start_c <- jags.data$start_c[4]
jags.data$end <- jags.data$end[4]
jags.data$end_c <- jags.data$end_c[4]
jags.data$pinit <- jags.data$pinit[4,]
jags.data$pinitS <- jags.data$pinitS[4]
jags.data$K <- 0

jags.data2 <- jags.data

jags.data2$marr <- marr
jags.data2$ns <- ns
jags.data2$rel <- rel
jags.data2$zero <- zero
jags.data2$ones <- ones

jags.data2$nyears <-jags.data2$n.occ

# saveRDS(jags.data2, file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Ilderton.rds")

##### Meadowlily ###########################################
Mmarr <- Capt_hist2 %>%
  filter(Population == "Meadowlily") %>%
  dplyr::select(-Population, -Individual, -Year, -first, -PopN)
Mmarr <- as.matrix(Mmarr)

marr <- marray(Mmarr)

# Bundle data
# Calculate the number of states
ns <- length(unique(as.numeric(Mmarr))) - 1

rel  <- rowSums(marr)
zero <- matrix(0, ns, ns)
ones <- diag(ns)

jags.data$n.occ <- 6
jags.data$Counts <- as.vector(t(jags.data$Counts[5,]))
jags.data$npop <- 5
jags.data$start <- jags.data$start[5]
jags.data$start_c <- jags.data$start_c[5]
jags.data$end <- jags.data$end[5]
jags.data$end_c <- jags.data$end_c[5]
jags.data$pinit <- jags.data$pinit[5,]
jags.data$pinitS <- jags.data$pinitS[5]
jags.data$K <- 0

jags.data2 <- jags.data

jags.data2$marr <- marr
jags.data2$ns <- ns
jags.data2$rel <- rel
jags.data2$zero <- zero
jags.data2$ones <- ones

jags.data2$nyears <-jags.data2$n.occ

# saveRDS(jags.data2, file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Meadowlily.rds")

###### RBG #################################################
Rmarr <- Capt_hist2 %>%
  filter(Population == "RBG") %>%
  dplyr::select(-Population, -Individual, -Year, -first, -PopN)
Rmarr <- as.matrix(Rmarr)

marr <- marray(Rmarr)

# Bundle data
# Calculate the number of states
ns <- length(unique(as.numeric(Rmarr))) - 1

rel  <- rowSums(marr)
zero <- matrix(0, ns, ns)
ones <- diag(ns)

jags.data$n.occ <- 6
jags.data$Counts <- as.vector(t(jags.data$Counts[6,]))
jags.data$npop <- 6
jags.data$start <- jags.data$start[6]
jags.data$start_c <- jags.data$start_c[6]
jags.data$end <- jags.data$end[6]
jags.data$end_c <- jags.data$end_c[6]
jags.data$pinit <- jags.data$pinit[6,]
jags.data$pinitS <- jags.data$pinitS[6]
jags.data$K <- 0

jags.data2 <- jags.data

jags.data2$marr <- marr
jags.data2$ns <- ns
jags.data2$rel <- rel
jags.data2$zero <- zero
jags.data2$ones <- ones

jags.data2$nyears <-jags.data2$n.occ

# saveRDS(jags.data2, file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_RBG.rds")
```

# JAGS Model

## JAGS settings
```{r Settings}
# LOad data
jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_25.rds")

# Initial values
inits <- function(){
  list(mean.logit.sy = runif(1, 0, 1), mean.logit.sa = runif(1, 0, 1))
}

# Parameters monitored
# These are the parameters that JAGS will keep track of, and what will be in
# the output
parameters <- c('N', 'Ntot', 'ann.growth.rate', 'mean.se', 'mean.germ',
                'mean.fl', 'sy', 'sj', 'sa', 'g1', 'g2', 'mean.sy',
                'mean.sj', 'mean.sa', 'mean.g1', 'mean.g2', 'mean.py',
                'mean.pj', 'mean.pa', 'sigma.sy', 'sigma.sj', 'sigma.sa',
                'sigma.g1', 'sigma.g2'
                )
```

## Bayfield

Create a text file of the JAGS model first
```{r Model}
# Model from Michael Schaub
cat(file = "IPM_Bayfield.txt", "

# Priors
model {

# Priors

mean.sy ~ dunif(0, 1)
mean.sj ~ dunif(0, 1)
mean.sa ~ dunif(0, 1)
mean.g1 ~ dunif(0, 1)
mean.g2 ~ dunif(0, 1)
mean.pj ~ dunif(0, 1)
mean.py ~ dunif(0, 1)
mean.pa ~ dunif(0, 1)


for (t in 1:(nyears-1+K)){
  #Seedlings
  sy[t] <- mean.sy
  # Small plants
  sj[t] <- mean.sj
  # Large plants
  sa[t] <- mean.sa

  g1[t] <- mean.g1
  g2[t] <- mean.g2
}

for (t in 1:(nyears-1)){
  py[t] <- mean.py       # recapture rate as seedling
  pj[t] <- mean.pj       # recapture rate as small plant
  pa[t] <- mean.pa       # recapture rate as large plant	
}


# Average germination rate
mean.germ ~ dbeta(105+1, (2025-105+1))   # instead of fixing it, use as a prior

for (t in 1:(nyears+K)){
  # Flowering
  fl[t,1] <- mean.fl[1]   # probability of flowering small plants
  fl[t,2] <- mean.fl[2]   # probability of flowering large plants
  
  se[t,1] <- mean.se[1]   # number seeds/plant for small plants
  se[t,2] <- mean.se[2]   # number seeds/plant for large plants

  germ[t] <- mean.germ    # Probability of seed germination  
}

mean.se[1] ~ dpois(0.461)  # number seeds/plant for small plants
mean.se[2] ~ dpois(144.29) # number seeds/plant for large plants 
mean.fl[1] ~ dunif(0, 1)   # probability of flowering small plants
mean.fl[2] ~ dunif(0, 1)   # probability of flowering small plants


# Population count data (state-space model) ##################################
# Model for the initial stage-specific population sizes: uniform priors

N[1,start_c] ~ dcat(piniS)  # Seedlings
N[2,start_c] ~ dcat(piniP)  # Small plants
N[3,start_c] ~ dcat(piniP)  # Large plants

# Process model over time: our model of population dynamics
# This is the life-cycle diagram

for (t in (start_c + 1):(end + K)){
  ## Seedlings
  N[1,t] ~ dpois(N[2,t-1] * fl[t-1,1] * se[t-1,1] * germ[t] + N[3,t-1] * fl[t-1,2] * se[t-1,2] * germ[t] + sy[t-1] * (1-g1[t-1]) * N[1,t-1])   

  ## Small plants
  N[2,t] ~ dpois(sy[t-1] * g1[t-1] * N[1,t-1] + sj[t-1] * (1-g2[t-1]) * N[2,t-1])

  # Large plants
  N[3,t] ~ dpois(sj[t-1] * g2[t-1] * N[2,t-1] + sa[t-1] * N[3,t-1])
}


# Observation model
# Counts exclude seedlings because they are very hard to find and count
for (t in start_c:end_c){
  Counts[t] ~ dpois(N[2,t] + N[3,t]) #### observed count is sum of stage classes
}

########################################################################
########################################################################

# Multistate data
  # Define state-transition and re-encounter probabilities
  for (t in 1:(nyears-1)){
    psi[1,t,1] <- sy[t] * (1-g1[t])
    psi[1,t,2] <- sy[t] * g1[t]
    psi[1,t,3] <- 0

    psi[2,t,1] <- 0
    psi[2,t,2] <- sj[t] * (1-g2[t])
    psi[2,t,3] <- sj[t] * g2[t]

    psi[3,t,1] <- 0
    psi[3,t,2] <- 0
    psi[3,t,3] <- sa[t]

    po[1,t] <- py[t]
    po[2,t] <- pj[t]
    po[3,t] <- pa[t]	
  }

  # From here onwards, no changes needed regardless of which model is fitted
  # Calculate probability of non-encounter (dq) and reshape the array for the encounter
  # probabilities
  for (t in 1:(nyears-1)){
    for (s in 1:ns){
      dp[s,t,s] <- po[s,t]
      dq[s,t,s] <- 1-po[s,t]
    } #s
    for (s in 1:(ns-1)){
      for (m in (s+1):ns){
        dp[s,t,m] <- 0
        dq[s,t,m] <- 0
      } #s
    } #m
    for (s in 2:ns){
      for (m in 1:(s-1)){
        dp[s,t,m] <- 0
        dq[s,t,m] <- 0
      } #s
    } #m
  } #t

  # Define the multinomial likelihood
  for (t in 1:((nyears-1)*ns)){
    marr[t,1:(nyears *ns-(ns-1))] ~ dmulti(pi[t,], rel[t])
  }

  # Define cell probabilities of the multistate m-array
  # Matrix U: product of probabilities of state-transition and non-encounter (needed because
  # there is no product function for matrix multiplication in JAGS)
  for (t in 1:(nyears-2)){
    U[(t-1)*ns+(1:ns),(t-1)*ns+(1:ns)] <- ones
    for (j in (t+1):(nyears-1)){
      U[(t-1)*ns+(1:ns),(j-1)*ns+(1:ns)] <- U[(t-1)*ns+(1:ns),(j-2)*ns+(1:ns)] %*% psi[,j-1,] %*% dq[,j-1,]
    } #j
  } #t
  U[(nyears-2)*ns+(1:ns), (nyears-2)*ns+(1:ns)] <- ones

  for (t in 1:(nyears-2)){
    # Diagonal
    pi[(t-1)*ns+(1:ns),(t-1)*ns+(1:ns)] <- U[(t-1)*ns+(1:ns),(t-1)*ns+(1:ns)] %*% psi[,t,] %*% dp[,t,]
    # Above main diagonal
    for (j in (t+1):(nyears-1)){
      pi[(t-1)*ns+(1:ns),(j-1)*ns+(1:ns)] <- U[(t-1)*ns+(1:ns),(j-1)*ns+(1:ns)] %*% psi[,j,] %*% dp[,j,]
    } #j
  } #t
  pi[(nyears-2)*ns+(1:ns),(nyears-2)*ns+(1:ns)] <- psi[,nyears-1,] %*% dp[,nyears-1,]

  # Below main diagonal
  for (t in 2:(nyears-1)){
    for (j in 1:(t-1)){
      pi[(t-1)*ns+(1:ns),(j-1)*ns+(1:ns)] <- zero
    } #j
  } #t

  # Last column: probability of non-recapture
  for (t in 1:((nyears-1)*ns)){
    pi[t,(nyears*ns-(ns-1))] <- 1-sum(pi[t,1:((nyears-1)*ns)])
  } #t  

##############################################################################
##############################################################################

# Derived parameters 
# Annual population growth rate
for (t in start_c:(end-1 + K)){
  ann.growth.rate[t] <- (N[1,t+1] + N[2, t+1] + N[3,t+1]) / (N[1,t] + N[2,t] + N[3,t] + 0.00000001)
}

# Total population size
for (t in start_c:(end+K)){
  Ntot[t] <- N[1,t] + N[2,t] + N[3,t] #### add pop size of each age class
}
} # end model

")

```

Run Model
```{r Run model-Bayfield}

jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Bayfield.rds")

# set projection period if needed
jags.data$K <- 50

jags.data$piniS <- dUnif(1, 500)
jags.data$piniP <- dUnif(5, 50)

# MCMC settings
ni <- 20000;  nb <- 10000;  nc <- 4; nt <- 10;  na <- 3000
# Runs for about 0.5 minutes

# Call JAGS, check convergence and summarize posteriors
startTime <- Sys.time()

out2 <- jags(jags.data, inits, parameters, "~/Research/Data/Population Models/IntegratedPopModel/IPM_Bayfield.txt",
             n.iter = ni, n.burnin = nb, n.chains = nc,
             n.thin = nt, n.adapt = na, parallel = T, set.seed(1))

endTime <- Sys.time(); endTime - startTime

# After the model runs, check the trace plots to confirm convergence
# The model I have now has pretty good convergence
# All Rhat values need to be less than 1.1

# traceplot(out2)

# Print the output of the model
 print(out2, 3)

#save model output
# saveRDS(out2, file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_B.rds")
```

## Denfield

```{r Run model-Denfield}
# Subset jags.data (loaded above) for just population 2
jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Denfield.rds")

# Set projection period 
jags.data$K <- 50

jags.data$piniS <- dUnif(1, 500)
jags.data$piniP <- dUnif(5, 50)

# Fix the m-array
jags.data$marr[1,2] <- 1
jags.data$marr[1,3] <- 0

jags.data$marr[4,5] <- 1
jags.data$marr[4,6] <- 0

# I think maybe I should use zero for the count in 2023, since there were no plants
jags.data$Counts[4] <- 0

# MCMC settings
ni <- 20000;  nb <- 10000;  nc <- 4; nt <- 10;  na <- 3000
# Runs for about 0.5 minutes

# Call JAGS, check convergence and summarize posteriors
startTime <- Sys.time()

out2 <- jags(jags.data, inits, parameters, "~/Research/Data/Population Models/IntegratedPopModel/IPM_Bayfield.txt",
             n.iter = ni, n.burnin = nb, n.chains = nc,
             n.thin = nt, n.adapt = na, parallel = T, set.seed(1))

endTime <- Sys.time(); endTime - startTime

# After the model runs, check the trace plots to confirm convergence
# The model I have now has pretty good convergence
# All Rhat values need to be less than 1.1

# traceplot(out2)

# Print the output of the model
 print(out2, 3)

#save model output
# saveRDS(out2, file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_D.rds")
```

## Fanshawe

```{r Run model-Fanshawe}

jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Fanshawe.rds")

jags.data$K <- 50

jags.data$piniS <- dUnif(1, 500)
jags.data$piniP <- dUnif(5, 50)

# MCMC settings

ni <- 20000;  nb <- 10000;  nc <- 4; nt <- 10;  na <- 3000
# Runs for about 0.5 minutes

# Call JAGS, check convergence and summarize posteriors
startTime <- Sys.time()

out2 <- jags(jags.data, inits, parameters, "~/Research/Data/Population Models/IntegratedPopModel/IPM_Bayfield.txt",
             n.iter = ni, n.burnin = nb, n.chains = nc,
             n.thin = nt, n.adapt = na, parallel = T, set.seed(1))

endTime <- Sys.time(); endTime - startTime

# After the model runs, check the trace plots to confirm convergence
# All Rhat values need to be less than 1.1

# traceplot(out2)

# Print the output of the model
 print(out2, 3)

#save model output
# saveRDS(out2, file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_F.rds")
```

## Ilderton

```{r Run model-Ilderton}
jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Ilderton.rds")

jags.data$K <- 50

jags.data$piniS <- dUnif(1, 500)
jags.data$piniP <- dUnif(500, 1000) # this is different compared to Bayfield, because of larger population size

## Fix m-array
jags.data$marr[7,8] <- 2
jags.data$marr[7,9] <- 0
jags.data$marr[13,14] <- 2
jags.data$marr[13,15] <- 0

# MCMC settings
ni <- 20000;  nb <- 10000;  nc <- 4; nt <- 10;  na <- 3000
# Runs for about 1.75 minutes

# Call JAGS, check convergence and summarize posteriors
startTime <- Sys.time()

out2 <- jags(jags.data, inits, parameters, "~/Research/Data/Population Models/IntegratedPopModel/IPM_Bayfield.txt",
             n.iter = ni, n.burnin = nb, n.chains = nc,
             n.thin = nt, n.adapt = na, parallel = T, set.seed(1))

endTime <- Sys.time(); endTime - startTime

# After the model runs, check the trace plots to confirm convergence
# The model I have now has pretty good convergence, but the vs2 is
# a little bad
# All Rhat values need to be less than 1.1

# traceplot(out2)

# Print the output of the model
 print(out2, 3)

#save model output
# saveRDS(out2, file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_I.rds")
```

## RBG

```{r Run model-RBG}
jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_RBG.rds")

jags.data$K <- 50

jags.data$piniS <- dUnif(1, 500)
jags.data$piniP <- dUnif(500, 1000) # this is different compared to Bayfield, because of larger population size

## Fix m-array
jags.data$marr[7,8] <- 1
jags.data$marr[7,9] <- 0
jags.data$marr[7,11] <- 1
jags.data$marr[7,12] <- 0
jags.data$marr[10,11] <- 2
jags.data$marr[10,12] <- 0

# MCMC settings
ni <- 20000;  nb <- 10000;  nc <- 4; nt <- 10;  na <- 3000
# Runs about 1.75 minutes

# Call JAGS, check convergence and summarize posteriors
startTime <- Sys.time()

out2 <- jags(jags.data, inits, parameters, "~/Research/Data/Population Models/IntegratedPopModel/IPM_Bayfield.txt",
             n.iter = ni, n.burnin = nb, n.chains = nc,
             n.thin = nt, n.adapt = na, parallel = T, set.seed(1))

endTime <- Sys.time(); endTime - startTime

# After the model runs, check the trace plots to confirm convergence
# The model I have now has pretty good convergence, but the vs2 is
# a little bad
# All Rhat values need to be less than 1.1

# traceplot(out2)

# Print the output of the model
 print(out2, 3)

#save model output
# saveRDS(out2, file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_R.rds")
```

# Use model output
```{r eval=FALSE, include=FALSE}
# Load IPM output
out2_B <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_B.rds")
out2_D <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_D.rds")
out2_F <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_F.rds")
out2_I <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_I.rds")
#out2_M <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_M.rds")
out2_R <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPopM_Output_R.rds")

```


## IPM vs IPM
Show Growth Rates for Both Model Types
For this, need bootstrapped growth rates for IProjM loaded as well.

Load Integral Projection Model output
```{r}
# Load IPM output
lamb_lists <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/IPM_Lambda.rds")

boot_lambda <- do.call(rbind, lamb_lists)

boot_lambda[,2] <- c(rep(1, 3000), rep(2, 3000), rep(3, 3000), rep(4, 3000))
boot_lambda[,3] <- rep(c(1, 2, 3), times = 4000)

names(boot_lambda) <- c("lambda", "Population", "Year_t")
str(boot_lambda)
boot_lambda <- boot_lambda %>%
  mutate(Population = case_when(Population == 1 ~ "Bayfield",
                                Population == 2 ~ "Fanshawe",
                                Population == 3 ~ "Ilderton",
                                Population == 4 ~ "RBG"),
         Year_t = case_when(Year_t == 1 ~ 2022,
                            Year_t == 2 ~ 2023,
                            Year_t == 3 ~ 2024))
boot_lambda$Population <- as.factor(boot_lambda$Population)
boot_lambda$Year_t <- as.factor(boot_lambda$Year_t)

# Mean and 95% CI
## Used for comparison with IPopM
data1 <- boot_lambda %>%
  group_by(Population, Year_t) %>%
  summarise(mean = mean(lambda),
            Q2.5 = quantile(lambda, probs = 0.025),
            Q97.5 = quantile(lambda, probs = 0.975))

p <- ggplot(data1, aes(x = Year_t, y = mean)) +
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), width = 0.1) +
    geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  geom_line() +
  facet_wrap(~Population) +
  geom_point() +
  theme_classic()
p
```


Organize and plot
```{r}
GrowthData_B <- (out2_B$q50$ann.growth.rate)
GrowthData_D <- (out2_D$q50$ann.growth.rate)
GrowthData_F <- (out2_F$q50$ann.growth.rate)
GrowthData_I <- (out2_I$q50$ann.growth.rate)
#GrowthData_M <- (out2_M$q50$ann.growth.rate)
GrowthData_R <- (out2_R$q50$ann.growth.rate)

GrowthData_Bl <- (out2_B$q2.5$ann.growth.rate)
GrowthData_Dl <- (out2_D$q2.5$ann.growth.rate)
GrowthData_Fl <- (out2_F$q2.5$ann.growth.rate)
GrowthData_Il <- (out2_I$q2.5$ann.growth.rate)
#GrowthData_Ml <- (out2_M$q2.5$ann.growth.rate)
GrowthData_Rl <- (out2_R$q2.5$ann.growth.rate)

GrowthData_Bu <- (out2_B$q97.5$ann.growth.rate)
GrowthData_Du <- (out2_D$q97.5$ann.growth.rate)
GrowthData_Fu <- (out2_F$q97.5$ann.growth.rate)
GrowthData_Iu <- (out2_I$q97.5$ann.growth.rate)
#GrowthData_Mu <- (out2_M$q97.5$ann.growth.rate)
GrowthData_Ru <- (out2_R$q97.5$ann.growth.rate)


Population <- rep("Bayfield", 55)
GrowthData_B <- data.frame(GrowthData_B, Population, GrowthData_Bl, GrowthData_Bu)
names(GrowthData_B) <- c("GrowthRate", "Population", "LCI","UCI")
GrowthData_B$Year <- seq(from = 2020, to = 2074, by = 1)

Population <- rep("Denfield", 55)
GrowthData_D <- data.frame(GrowthData_D, Population, GrowthData_Dl, GrowthData_Du)
names(GrowthData_D) <- c("GrowthRate", "Population", "LCI","UCI")
GrowthData_D$Year <- seq(from = 2020, to = 2074, by = 1)

Population <- rep("Fanshawe", 55)
GrowthData_F <- data.frame(GrowthData_F, Population, GrowthData_Fl, GrowthData_Fu)
names(GrowthData_F) <- c("GrowthRate", "Population", "LCI","UCI")
GrowthData_F$Year <- seq(from = 2020, to = 2074, by = 1)

Population <- rep("Ilderton", 55)
GrowthData_I <- data.frame(GrowthData_I, Population, GrowthData_Il, GrowthData_Iu)
names(GrowthData_I) <- c("GrowthRate", "Population", "LCI","UCI")
GrowthData_I$Year <- seq(from = 2020, to = 2074, by = 1)


#Population <- rep("Meadowlily", 53)
#GrowthData_M <- data.frame(GrowthData_M, Population, GrowthData_Ml, GrowthData_Mu)
#names(GrowthData_M) <- c("GrowthRate", "Population", "LCI","UCI")
#GrowthData_M$Year <- seq(from = 2020, to = 2072, by = 1)
#GrowthData_M <- GrowthData_M %>%
#  as.data.frame() %>%
#  add_row(GrowthRate = NA, Population = "Meadowlily", LCI = NA, UCI = NA, Year = 2073) %>%
#    add_row(GrowthRate = NA, Population = "Meadowlily", LCI = NA, UCI = NA, Year = 2074) %>%
#    add_row(GrowthRate = NA, Population = "Meadowlily", LCI = NA, UCI = NA, Year = 2075) %>%
#    add_row(GrowthRate = NA, Population = "Meadowlily", LCI = NA, UCI = NA, Year = 2076)

Population  <- rep("RBG", 55)
GrowthData_R <- data.frame(GrowthData_R, Population, GrowthData_Rl, GrowthData_Ru)
names(GrowthData_R) <- c("GrowthRate", "Population", "LCI","UCI")
GrowthData_R$Year <- seq(from = 2020, to = 2074, by = 1)

GrowthData <- rbind(GrowthData_B, GrowthData_D, GrowthData_F, GrowthData_I, GrowthData_R)

PopSize <- as.data.frame(GrowthData)
PopSize <- PopSize %>%
  filter(!is.na(GrowthRate))


##########################################################
# both!!

names(data1) <- c("Population", "Year", "GrowthRate", "LCI", "UCI")
data1$ModelType <- "IProjM"
data1$Year <- as.numeric(data1$Year)
data1 <- data1 %>%
  mutate(Year = case_when(Year == 1 ~ 2022,
                          Year == 2 ~ 2023,
                          Year == 3 ~ 2024))

GrowthData$ModelType <- "IPopM"

data2 <- rbind(data1, GrowthData)
data2 <- data2 %>%
  filter(Year < 2025)

data2$ModelType <- factor(data2$ModelType, levels = c("IPopM","IProjM"))

data2 <- data2 %>%
  mutate(Population = case_when(Population == "RBG" ~ "RBG - Transplanted",
                                TRUE ~ Population))

data2 <- data2 %>%
 mutate(GrowthRate = case_when(Population == "Meadowlily" & Year > 2022 ~ NA,
                               TRUE ~ GrowthRate),
        LCI = case_when(Population == "Meadowlily" & Year > 2022 ~ NA,
                               TRUE ~ LCI),
        UCI = case_when(Population == "Meadowlily" & Year > 2022 ~ NA,
                               TRUE ~ UCI))

# Jitter the points
data3 <- data2 %>%
  mutate(Year = case_when(ModelType == "IPopM" ~ Year + 0.25,
                        TRUE ~ Year))
```

Add count-based
```{r}
Counts <- read.csv("~/Research/Data/Count_Data.csv")

Wood_Poppy_Vertical <- read.csv("~/Research/Data/IPM_Formatted_STYL/Wood_Poppy_Vertical.csv")

Counts <- Counts %>%
  filter(Species == "Stylophorum") %>%
#  mutate(Count = case_when(Population == "Ilderton" & Year > 2021 ~ NA_real_,
#                           TRUE ~ Count)) %>%
  dplyr::select(-Species)


#Quadrats <- Wood_Poppy_Vertical %>%
#  filter(Population == "Ilderton") %>%
#  filter(Status == "Alive") %>%
#  group_by(Population, Year) %>%
#  summarise(Count = n())


CountLambda <- as.data.frame(rbind(Counts))#, Quadrats))

CountLambda <- CountLambda %>%
  filter(!is.na(Count)) %>%
  arrange(Population, Year) %>%
  group_by(Population) %>%
  mutate(GrowthRate = Count / lag(Count)) %>%
  mutate(Year = Year - 1.25) %>%
  ungroup() %>%
  mutate(ModelType = "Count") %>%
  mutate(LCI = NA_real_,
         UCI = NA_real_) %>%
  mutate(Population = case_when(Population == "RBG" ~ "RBG - Transplanted",
                                TRUE ~ Population)) %>%
  dplyr::select(-Count)

# Ilderton 2021-202
#CountLambda[21, 3] <- NA
# Denfield
CountLambda[12, 3] <- NA
CountLambda[11, 3] <- NA

# Fix RBG count to be the average (so the nth root of the overall)
CountLambda[31,3] <- CountLambda[31,3]^(1/3)


head(CountLambda)

AllData <- rbind(data3, CountLambda)
```


Plot
```{r}
# make labels
PopFacet <- c("Bayfield" = "(a) Bayfield", "Denfield" = "(b) Denfield", "Fanshawe" = "(c) Fanshawe", "Ilderton" = "(d) Ilderton", "RBG - Transplanted" = "(e) RBG")

p <-ggplot(AllData %>%
             filter(Population != "Meadowlily") %>%
             filter(Year > 2019 & Year < 2024.75) %>%
             
             mutate(Population = case_when(Population == "Bayfield" ~ "(a) Bayfield",
                                           Population == "Denfield" ~ "(b) Denfield",
                                           Population == "Fanshawe" ~ "(c) Fanshawe",
                                           Population == "Ilderton" ~ "(d) Ilderton",
                                           Population == "RBG - Transplanted" ~ "(e) RBG")), aes(x = Year, y = GrowthRate, color = ModelType, shape = ModelType)) +
  geom_errorbar(aes(ymin = LCI, ymax = UCI), width = 0.6) +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  scale_color_manual(values = c("gray25", "#6A0F8E","#4ECDC4"), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  scale_shape_manual(values = c(8,16,17), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  # show every x axis value
  scale_x_continuous(breaks = seq(2020, 2024, by = 1)) +

  #geom_line() +
  #facet_wrap(~Population, labeller = as_labeller(PopFacet)) +
  geom_point(size = 3) +
  theme_classic() +
  theme(legend.position = c(0.84,0.1)) +#c(0.13, 0.1)) +#c(0.15, 0.1)) + #c(0.15, 0.92)) +
  theme(strip.background = element_blank()) +
  theme(strip.text = element_text(hjust = 0)) +
  # adds axes to each facet
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  
  
  theme(strip.text = element_text( size = 14, color = "black")) +
  theme(axis.text = element_text(size = 14)) +
  theme(axis.title = element_text(size = 14)) +
  theme(legend.title = element_text(size=14), #change legend title font size
        legend.text = element_text(size=14))
#p

# add back the tick marks on all plots
p1 <- p +  facet_wrap2(vars(Population), axes = "all", remove_labels = "all")
p1

#############################
# Make with multiple panels
p <- ggplot(AllData %>%
              filter(Population == "Bayfield"), aes(x = Year, y = GrowthRate, color = ModelType, shape = ModelType)) +
  geom_errorbar(aes(ymin = LCI, ymax = UCI), width = 0.6) +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  scale_color_manual(values = c("lightgray", "#6A0F8E","#4ECDC4"), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  scale_shape_manual(values = c(8,16,17), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  #geom_line() +
  #facet_wrap(~Population) + #, scales = "free_y") +
  geom_point(size = 3) +
  theme_classic() +
  scale_y_continuous(limits = c(0, 1.8)) +
  theme(legend.position = "none") + #c(0.15, 0.92)) +
  theme(axis.text = element_text(size = 14)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

p


p2 <- ggplot(AllData %>%
               filter(Population == "Denfield"), aes(x = Year, y = GrowthRate, color = ModelType, shape = ModelType)) +
  geom_errorbar(aes(ymin = LCI, ymax = UCI), width = 0.6) +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  scale_color_manual(values = c("lightgray", "#6A0F8E","#4ECDC4"), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  scale_shape_manual(values = c(8,16,17), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  
  #geom_line() +
  #facet_wrap(~Population) + #, scales = "free_y") +
  geom_point(size = 3) +
  theme_classic() +
  scale_y_continuous(limits = c(0, 1.8)) +
  theme(legend.position = "none") + #c(0.15, 0.92)) +
  theme(axis.text = element_text(size = 14)) +
  theme(axis.title = element_blank(),
        axis.text=element_blank())
p3 <- ggplot(AllData %>%
               filter(Population == "Fanshawe"), aes(x = Year, y = GrowthRate, color = ModelType, shape = ModelType)) +
  geom_errorbar(aes(ymin = LCI, ymax = UCI), width = 0.6) +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  scale_color_manual(values = c("lightgray", "#6A0F8E","#4ECDC4"), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  scale_shape_manual(values = c(8,16,17), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  
  #geom_line() +
  #facet_wrap(~Population) + #, scales = "free_y") +
  geom_point(size = 3) +
  theme_classic() +
  scale_y_continuous(limits = c(0, 1.8)) +
  theme(legend.position = "none") + #c(0.15, 0.92)) +
  theme(axis.text = element_text(size = 14)) +
  theme(axis.title.y = element_blank(),
        axis.text.y=element_blank())
p4 <- ggplot(AllData %>%
               filter(Population == "Ilderton"), aes(x = Year, y = GrowthRate, color = ModelType, shape = ModelType)) +
  geom_errorbar(aes(ymin = LCI, ymax = UCI), width = 0.6) +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  scale_color_manual(values = c("lightgray", "#6A0F8E","#4ECDC4"), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  scale_shape_manual(values = c(8,16,17), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  
  #geom_line() +
  #facet_wrap(~Population) + #, scales = "free_y") +
  geom_point(size = 3) +
  theme_classic() +
  scale_y_continuous(limits = c(0, 1.8)) +
  theme(legend.position = c(0.45, 0.2)) + #c(0.15, 0.92)) +
  theme(axis.text = element_text(size = 14)) +
  theme(legend.title = element_text(size=10), #change legend title font size
        legend.text = element_text(size=10))
p5 <- ggplot(AllData %>%
               filter(Population == "RBG - Transplanted"), aes(x = Year, y = GrowthRate, color = ModelType, shape = ModelType)) +
  geom_errorbar(aes(ymin = LCI, ymax = UCI), width = 0.6) +
  geom_hline(yintercept = 1, col = "darkred", linetype = 2) +
  labs(x = "Year t", y = expression(paste("Growth rate (", lambda, ")"))) +
  scale_color_manual(values = c("lightgray", "#6A0F8E","#4ECDC4"), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  scale_shape_manual(values = c(8,16,17), name = "Model Type", labels = c("Counts", "Integrated Population Model", "Integral Projection Model")) +
  
  #geom_line() +
  #facet_wrap(~Population) + #, scales = "free_y") +
  geom_point(size = 3) +
  theme_classic() +
  scale_y_continuous(limits = c(0, 1.8)) +
  theme(legend.position = "none") + #c(0.15, 0.92)) +
  theme(axis.text = element_text(size = 14)) +
  theme(axis.title.y = element_blank(),
        axis.text.y=element_blank())


Graphs <- ggarrange(p,  p2,  p3,  p4, p5,
                      labels = c("(a)", "(b)", "(c)", "(d)", "(e)"),
                    #widths = c(1,1,1,1),
                      font.label = list(size = 11), 
                      hjust = c(-3.5, -3.5, -3.5, -3.5, -3.5),
                      #hjust = c(-1.5,-0.2,-0.2, -1.5, -0.2, -0.2),
                      ncol = 3, nrow = 2, align = "hv"
                    )
Graphs


# Save graph
# ggsave(plot = p1, dpi = 600,
#       filename = "Wood_Poppy_IProjM_IPopM_Lambda.png",
#       path = "C:/Users/ComputerName/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs",
#       width = 9, height = 5, units = "in")
# ggsave(plot = p1, dpi = 600,
#       filename = "Wood_Poppy_IProjM_IPopM_Lambda.pdf",
#       path = "C:/Users/ComputerName/Documents/uLethbridge/Ch_1_Wood-poppy/Graphs",
#       width = 9, height = 5, units = "in")


```




# Recapture rate
```{r}
# RBG
jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_RBG.rds")

## Fix m-array
jags.data$marr[7,8] <- 1
jags.data$marr[7,9] <- 0
jags.data$marr[7,11] <- 1
jags.data$marr[7,12] <- 0
jags.data$marr[10,11] <- 2
jags.data$marr[10,12] <- 0

sum(jags.data$marr[,1:15])
sum(jags.data$marr[,16])

306/(306 + 174) * 100

# Ilderton
jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Ilderton.rds")

## Fix m-array
jags.data$marr[7,8] <- 2
jags.data$marr[7,9] <- 0
jags.data$marr[13,14] <- 2
jags.data$marr[13,15] <- 0

sum(jags.data$marr[,1:15])
sum(jags.data$marr[,16])

399/(399 + 129) * 100


# Bayfield
jags.data <- readRDS(file = "~/Research/Data/IPM_Formatted_STYL/Output/JAGSdata_Bayfield.rds")

sum(jags.data$marr[,1:15])
sum(jags.data$marr[,16])

441/(441+66) * 100
```

# Vital Rates
```{r}

# ann.growth.rate
mean(out2_B$q50$ann.growth.rate[1:5])
mean(out2_D$q50$ann.growth.rate[1:5])
mean(out2_F$q50$ann.growth.rate[1:5])
mean(out2_I$q50$ann.growth.rate[1:5])
mean(out2_R$q50$ann.growth.rate[2:5])

mean(out2_B$sd$ann.growth.rate[1:5])
mean(out2_D$sd$ann.growth.rate[1:5])
mean(out2_F$sd$ann.growth.rate[1:5])
mean(out2_I$sd$ann.growth.rate[1:5])
mean(out2_R$sd$ann.growth.rate[2:5])

mean(out2_B$q2.5$ann.growth.rate[1:5])
mean(out2_D$q2.5$ann.growth.rate[1:5])
mean(out2_F$q2.5$ann.growth.rate[1:5])
mean(out2_I$q2.5$ann.growth.rate[1:5])
mean(out2_R$q2.5$ann.growth.rate[2:5])

mean(out2_B$q97.5$ann.growth.rate[1:5])
mean(out2_D$q97.5$ann.growth.rate[1:5])
mean(out2_F$q97.5$ann.growth.rate[1:5])
mean(out2_I$q97.5$ann.growth.rate[1:5])
mean(out2_R$q97.5$ann.growth.rate[2:5])




# SA
mean(out2_B$q50$sa[1:5])
mean(out2_D$q50$sa[1:5])
mean(out2_F$q50$sa[1:5])
mean(out2_I$q50$sa[1:5])
mean(out2_R$q50$sa[1:5])

mean(out2_B$sd$sa[1:5])
mean(out2_D$sd$sa[1:5])
mean(out2_F$sd$sa[1:5])
mean(out2_I$sd$sa[1:5])
mean(out2_R$sd$sa[1:5])

mean(out2_B$q2.5$sa[1:5])
mean(out2_D$q2.5$sa[1:5])
mean(out2_F$q2.5$sa[1:5])
mean(out2_I$q2.5$sa[1:5])
mean(out2_R$q2.5$sa[1:5])

mean(out2_B$q97.5$sa[1:5])
mean(out2_D$q97.5$sa[1:5])
mean(out2_F$q97.5$sa[1:5])
mean(out2_I$q97.5$sa[1:5])
mean(out2_R$q97.5$sa[1:5])



# SJ
mean(out2_B$q50$sj[1:5])
mean(out2_D$q50$sj[1:5])
mean(out2_F$q50$sj[1:5])
mean(out2_I$q50$sj[1:5])
mean(out2_R$q50$sj[1:5])

mean(out2_B$sd$sj[1:5])
mean(out2_D$sd$sj[1:5])
mean(out2_F$sd$sj[1:5])
mean(out2_I$sd$sj[1:5])
mean(out2_R$sd$sj[1:5])

mean(out2_B$q2.5$sj[1:5])
mean(out2_D$q2.5$sj[1:5])
mean(out2_F$q2.5$sj[1:5])
mean(out2_I$q2.5$sj[1:5])
mean(out2_R$q2.5$sj[1:5])

mean(out2_B$q97.5$sj[1:5])
mean(out2_D$q97.5$sj[1:5])
mean(out2_F$q97.5$sj[1:5])
mean(out2_I$q97.5$sj[1:5])
mean(out2_R$q97.5$sj[1:5])


# SY
mean(out2_B$q50$sy[1:5])
mean(out2_D$q50$sy[1:5])
mean(out2_F$q50$sy[1:5])
mean(out2_I$q50$sy[1:5])
mean(out2_R$q50$sy[1:5])

mean(out2_B$sd$sy[1:5])
mean(out2_D$sd$sy[1:5])
mean(out2_F$sd$sy[1:5])
mean(out2_I$sd$sy[1:5])
mean(out2_R$sd$sy[1:5])

mean(out2_B$q2.5$sy[1:5])
mean(out2_D$q2.5$sy[1:5])
mean(out2_F$q2.5$sy[1:5])
mean(out2_I$q2.5$sy[1:5])
mean(out2_R$q2.5$sy[1:5])

mean(out2_B$q97.5$sy[1:5])
mean(out2_D$q97.5$sy[1:5])
mean(out2_F$q97.5$sy[1:5])
mean(out2_I$q97.5$sy[1:5])
mean(out2_R$q97.5$sy[1:5])




```

